### 2019-12-27 lloveras_2020_Albert.tex
Ediciones por ALBERT del VIERNES 27

%% Template.tex; Solar Physics
%% 
\documentclass[namedreferences]{solarphysics}
%
% spr-sola-addons available options:
%  hyperref      -- loads hyperref.sty with options (pdfborder={0 0 0 },urlcolor=blue,breaklinks)
%  nonatbib      -- do not load natbib.sty (style loads it by default)
%  solaromanenum -- makes enumerated list with roman numerals and a single right-bracket
%  linksfromyear -- puts a link on a year citation (hyperref must be loaded). Loaded by default
%  nolinksfromyear -- suppress  linksfromyear
%  optionalrh    -- for optional running title/author
%  showbiblabels -- to show bibitem label at end of bibitem (via \endbibitem command)
%
\usepackage[hyperref,optionalrh,solaromanenum]{spr-sola-addons} % For Solar Physics 
%\usepackage{epsfig}                     % For eps figures, old commands
\usepackage{graphicx}                    % For eps figures, newer & more powerfull
%\usepackage{courier}                    % Change the \texttt command to courier style
\usepackage{amssymb}                     % useful mathematical symbols
\usepackage{color}                       % For color text: \color command
\usepackage{breakurl}                    % For breaking URLs easily trough lines
\def\UrlFont{\sf}                        % define the fonts for the URLs
\usepackage{comment}
\usepackage{setspace}
\usepackage[utf8]{inputenc}

%% Local definitions
%% please place your own definitions here and don't use \def but
%% \newcommand{}{} or 
%% \renewcommand{}{} if it is already defined in LaTeX
\renewcommand{\deg}{$^\circ$}
\newcommand{\mdeg}{^\circ}
\newcommand{\rsun}{R{$_\odot$}}
\renewcommand{\l}{\lambda_{\rm N}}%\lambda_
\newcommand{\mrsun}{{\rm R_\odot}}
\newcommand{\med}{{\rm Md}}
\newcommand{\avgTe}{\left<\Tm\right>}
\newcommand{\MK}{{\rm MK}}
\newcommand{\cm}{{\rm cm}}
\newcommand{\cminvc}{\cm^{-3}}
\newcommand{\cminvs}{\cm^{-2}}
\newcommand{\erg}{{\rm erg}}
\newcommand{\s}{{\rm s}}
\newcommand{\sdev}{{\rm SD}}
\newcommand{\mean}{{\rm Mn}}
\newcommand{\deltat}{$\delta$}
\newcommand{\LDEM}{{\rm LDEM}}
\newcommand{\FBE}{{\rm FBE}}
\newcommand{\TRF}{{\rm TRF}}
\newcommand{\lN}{\lambda_N}
\newcommand{\NCB}{N_{\rm CB}}
\newcommand{\Nqi}{N_{\rm QI}}

\newcommand{\dr}{\triangledown_r}
\newcommand{\er}{\mathbf{e}_r}
\newcommand{\Te}{T_{\rm e}}
\newcommand{\Tm}{T_{\rm m}}
\newcommand{\aTm}{\left<\Tm\right>}
\newcommand{\Tmi}{T_{{\rm m},i}}
\newcommand{\Ne}{N_{\rm e}}
\newcommand{\Nm}{N_{\rm m}}
\newcommand{\Nsqmi}{N^2_{{\rm m},i}}
\newcommand{\Nsqm}{N^2_{{\rm m}}}
\newcommand{\rhoTr}{\rho(T,r)}
\newcommand{\Tefit}{T_{e,{\rm fit}}}
\newcommand{\sqravgNi}{\sqrt{\Nsqmi}}
\newcommand{\sqravgN}{\sqrt{\Nsqm}}
\newcommand{\Tc}{T_{\rm c}}

\newcommand{\BibTeX}{\textsc{Bib}\TeX}
\newcommand{\etal}{{\it et al.}}

\newcommand{\Pl}{\texttt{+}}
\newcommand{\Mi}{\texttt{-}}

% Definitions for the journal names
\newcommand{\adv}{    {\it Adv. Space Res.}} 
\newcommand{\annG}{   {\it Ann. Geophys.}} 
\newcommand{\aap}{    {\it Astron. Astrophys.}}
\newcommand{\aaps}{   {\it Astron. Astrophys. Suppl.}}
\newcommand{\aapr}{   {\it Astron. Astrophys. Rev.}}
\newcommand{\ag}{     {\it Ann. Geophys.}}
\newcommand{\aj}{     {\it Astron. J.}} 
\newcommand{\apj}{    {\it Astrophys. J.}}
\newcommand{\apjs}{   {\it Astrophys. J. Suppl.}}
\newcommand{\apjl}{   {\it Astrophys. J. Lett.}}
\newcommand{\apss}{   {\it Astrophys. Space Sci.}} 
\newcommand{\cjaa}{   {\it Chin. J. Astron. Astrophys.}} 
\newcommand{\gafd}{   {\it Geophys. Astrophys. Fluid Dyn.}}
\newcommand{\grl}{    {\it Geophys. Res. Lett.}}
\newcommand{\ijga}{   {\it Int. J. Geomagn. Aeron.}}
\newcommand{\jastp}{  {\it J. Atmos. Solar-Terr. Phys.}} 
\newcommand{\jgr}{    {\it J. Geophys. Res.}}
\newcommand{\mnras}{  {\it Mon. Not. Roy. Astron. Soc.}}
\newcommand{\nat}{    {\it Nature}}
\newcommand{\pasp}{   {\it Pub. Astron. Soc. Pac.}}
\newcommand{\pasj}{   {\it Pub. Astron. Soc. Japan}}
\newcommand{\pre}{    {\it Phys. Rev. E}}
\newcommand{\solphys}{{\it Solar Phys.}}
\newcommand{\sovast}{ {\it Soviet  Astron.}} 
\newcommand{\ssr}{    {\it Space Sci. Rev.}}

%---------REMOVE THESE OPTIONS AND DEFs BEFORE SUMBISSION---------------------------------------
\usepackage[usenames,dvipsnames]{xcolor} % Las opciones: [USENAMES,DVIPSNAMES] 
                                         % ponen a disposición los colores ejemplificados aquí:
                                         % https://en.wikibooks.org/wiki/LaTeX/Colors
\def\diego#1{\textcolor{red}{#1}}
\def\albert#1{\textcolor{orange}{#1}}
\def\fede#1{\textcolor{ForestGreen}{#1}}
\def\temp#1{\textcolor{gray}{#1}}
\def\notebyalbert#1{\textcolor{blue}{NOTE: #1}}
%-----------------------------------------------------------------------------------------------

\begin{document}

\begin{article}

\begin{opening}

\title{Thermodynamic Structure of the Solar Corona: Tomographic Reconstructions and MHD Modeling}

\author[addressref={aff1},corref,email={dlloveras@iafe.uba.ar}]{\inits{D.G.}\fnm{Diego G.}~\lnm{Lloveras}\orcid{0000-0003-1402-0398}}

\author[addressref={aff1,aff2},corref,email={albert@iafe.uba.ar}]{\inits{A.M.}\fnm{Alberto M.}~\lnm{V\'asquez}\orcid{0000-0003-3401-6409}}

\author[addressref={aff1,aff3},corref,email={federico@iafe.uba.ar}]{\inits{F.A.}\fnm{Federico A.}~\lnm{Nuevo}\orcid{0000-0003-2355-5853}}

\author[addressref={aff1,aff2},corref,email={cmaccormack@iafe.uba.ar}]{\inits{C.}\fnm{Cecilia}~\lnm{Mac Cormack}\orcid{0000-0003-1173-503X}}

\author[addressref={aff4},corref,email={nishthas@umich.edu}]{\inits{N.}\fnm{Nishtha}~\lnm{Sachdeva}\orcid{0000-0001-9114-6133}}

\author[addressref={aff4},corref,email={chipm@umich.edu}]{\inits{W.}\fnm{Ward}~\lnm{Manchester IV}\orcid{0000-0003-0472-9408}}

\author[addressref={aff4},corref,email={bartvand@umich.edu}]{\inits{B.}\fnm{Bartholomeus}~\lnm{Van der Holst}\orcid{0000-0001-5260-3944}}

\author[addressref={aff4},corref,email={rfrazin@umich.edu}]{\inits{R.A.}\fnm{Richard A.}~\lnm{Frazin}\orcid{0000-0002-0281-7677}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Runningheads
%
\runningauthor{D.G.~Lloveras~\etal}
%\runningtitle{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Affilations 
%% id shold be the same with \author addressref value.
%\address[id={}]{}
\address[id=aff1]{Instituto de Astronom\'{\i}a y F\'{\i}sica del Espacio (IAFE), CONICET-UBA, CC 67 - Suc 28,  (C1428ZAA) Ciudad Aut\'onoma de Buenos Aires, Argentina}

\address[id=aff2]{Universidad Nacional de Tres de Febrero (UNTREF). Departamento de Ciencia y Tecnolog\'{\i}a, Sáenz Peña, Argentina.}

\address[id=aff3]{Ciclo Básico Común (CBC), Universidad de Buenos Aires (UBA), Buenos Aires, Argentina}

\address[id=aff4]{Department of Climate and Space Sciences and Engineering (CLaSP), University of Michigan, 2455 Hayward Street, Ann Arbor, MI 48109-2143, USA}

\begin{abstract}
Observational {techniques} play an essential role in advancing our understanding of the physics of the solar corona. They provide validation data {for} three-dimensional (3D) magnetohydrodynamic (MHD) models of the solar atmosphere, {key to improve} their space weather forecasting capabilities. Solar rotational tomography (SRT) is currently the sole observational technique that provides an empirical 3D description of some {of the fundamental} plasma parameters of the solar corona at a global scale. Based on EUV data of space borne instruments, SRT allows constructing 3D maps of the coronal electron density and temperature at heliocentric heights below 1.25 Rsun. We carry out a study of the corona combining tomographic reconstructions with MHD simulations {using} the latest version of the Alfv\'en Wave Solar Model (AWSoM) of the Space Weather Modeling Framework (SWMF). Target rotations were selected from the solar minimum between solar cycles (SC) 23 and 24 and the current declining phase of solar cycle 24. Tomographic reconstructions and results of the model are analyzed in distinct coronal magnetic structures. We study magnetically closed structures associated with the equatorial streamer belt, as well as magnetically open regions enclosing it. We report on the tomographic results in the different structures, their implications for the physics of the solar corona, and the {current} capability of the AWSoM model to reproduce the tomographic reconstructions in different regions.

\end{abstract}

\keywords{Solar Cycle, Observations; Corona,E; Corona, Structures}

\end{opening}

\section{Introduction}\label{intro}

Being the place where the solar atmosphere is heated and the solar wind accelerated, and where impulsive events such as solar flares and coronal mass ejections are energized, observation and modeling of the solar corona are tasks of great relevance to improve our understanding of the Sun-Earth environment. To advance our knowledge of the physics of the solar corona, as well as to improve its three-dimensional (3D) models, information derived from observational data plays a key role. Solar rotational tomography is currently the sole observational technique able to provide a quantitative empirical description of the 3D distribution of some fundamental plasma parameters of the solar corona at a global scale.

\noindent\notebyalbert{Here we will put in a couple of paragraphs summarizing what SRT and DEMT are, including references and defining the accronyms in the process. We should refer to minima in particular. We should mention the instruments and missions we use as data for DEMT, and define their accronyms too. Also, a couple of paragraphs on the AWSoM model and the SWMF, including references and defining the accronyms in the process.}

\section{Methodology}\label{meto}

\subsection{DEMT reconstructions}\label{demt}

\noindent\notebyalbert{This section will summarize DEMT's main aspects and equations, only those needed to understand the paper. Nothing not needed to read this paper is provided. Also, no derivation of equations are provided. References are given so that the reader can dig deeper if so desires. Points to cover are:
\begin{itemize}
  \item Selected targets, instruments used. One paragraph.
  \item Characteristics of the data (specific time series, dates, bands, etc.), SolarSoft software version used for their processing. One paragraph.
  \item Technical aspects of DEMT (grid, regularization, off-limb data is only used) and how these differ from previous versions (rereferences). One paragraph.
  \item What is FBE, what is LDEM, what is R. Why we care about these three quantities and what is their physical meaning should be clear. Then the final products $\Nm$ and $\Tm$, and what is their physical meaning.
\end{itemize}
}

To perform the comparison two specific rotations were selected, CR-2082 (2009, 05 April through 03 May), a deep minimum period between solar cycles (SCs) 23 and 24 characterized by virtually no ARs, and CR-2208 (2018, 02 September through 29 September), a rotation during the early declining fase of SC 24.

To determine the thermodynamical structure of both rotations, the DEMT technique was applied to respective time series of EUV images covering each first half rotation. To study CR-2082 and CR-2208 data taken by the EUVI-B/STEREO and AIA/SDO instruments was used, respectively.

EUV images are affected by instrumental stray light contamination, which can be corrected by deconvolution of the point spread function (PSF). While in the case of AIA images there is no reliable determination of its PSF, it has been well determined for EUVI-B by \citet{shearer_2012}. For this article, all EUVI images were stray-light corrected. For the details of impact in $\Nm$ and $\Tm$, refer to \citet{lloveras_ba2017}.

In DEMT, the low corona in the height range $1.0$ to $1.25\,\mrsun$ is discretized in a spherical computational grid. The size of the tomographic cell (or voxel) is set to 0.01 \rsun \ in the radial direction, while is set 2\deg\ in both the latitudinal and longitudinal directions. We carried out a reconstruction using half of a solar rotation blocking the lines-of-sight comming from the disk (i.e. only using off-limb data).

Using the high cadence of data available, one image every 10 minutes were used. The selected images were averaged in 6-hour intervals which, given the angular resolution, is the cadence needed to fully constrain the inversion problem, for a total of about 55 images to cover a half solar rotation. For both rotations, the first 14 days of data were used.

\diego{ DONDE QUEDA MEJOR COMENTAR ESTA MEJORA?
Unlike in previous works, a 3D regularization scheme was applied, thus spurious artifacts (nduced by the previous longitude latitude regularization schemei) were minimized in the first two heights of the tomographic grid.}

We refer the reader to \citet{frazin_2009} for a detailed description of the technique, and to \citet{vasquez_2016} for a recent review on all published work based on it. Here we summarize the key points. The technique involves two consecutive procedures.

In a first step, the time series of EUV images is used to solve a SRT problem, for each EUV band independently. As a result, the 3D distribution of the so called {filter band emissivity} (FBE) is determined for each band separately. The FBE, {an emissivity-type quantity}, is defined as the wavelength integral of the coronal EUV spectral emissivity and the telescope's passband function of each EUV channel. Line-of-sight (LOS) integration of the FBE provides synthetic images that can be quantitatively compared to the real data in the time series. To find the FBE, the tomographic problem is posed a global optimization problem in which the quadratic norm of the difference between all pairs of synthetic and real images is minimized. Spatial regularization terms are also included in the objective function as to minimize high frequency spurious artifacts that are due to the specific characteristics of the SRT sparse projection matrix \citep{frazin_2009,frazin_2000}. {Due to optical depth issues that may affect LOSs} passing very close to the limb, and to the decay of intensity with radii, the 3D maps produced by DEMT analysis typically cover the height range 1.02 to 1.23 \rsun \ \citep{frazin_2009}.

\diego{COMO INTRODUZCO EL NIVEL DE REGULARIZACION? For each instrument, the DEMT analysis involves a cross-validation study to determine the optimal level of regularization.}

{In a second step, the FBE values obtained for all bands in each voxel are used to constrain the determination of a {local differential emission measure} (LDEM) distribution, which describes the temperature distribution of the electron plasma contained in each individual tomographic grid voxel. Specifically, at each tomographic voxel $i$, the FBE of the band $k$ is related to the LDEM of the voxel according to}

\begin{equation}\label{fbeldem}
\FBE_i^{(k)}  = \int\,\textrm{d}T\,\LDEM_i(T)\,{\TRF}^{(k)}(T), \ \ \ {k=1,...,K}
\end{equation}

{
Due to unresolved coronal dynamics, tomographic reconstructions exhibit negative values of the reconstructed FBE, or zero when the solution is constrained to positive values \citep{frazin_2000,frazin_2009}. These non-reconstructed voxels are indicated as black voxels in the Carrington maps of the reconstructed DEMT reults in Section \ref{resu}.
}

Once the LDEM is determined at each voxel, the average squared electron density $\langle \Ne^2 \rangle$ and the electron mean temperature $\Tm$ in the voxel can be computed by taking its zeroth and first moments over temperature. More specifically, at the $i$-th voxel,

{
\begin{eqnarray}\label{momento1}
 \Nsqmi = \langle \Ne^2 \rangle_i &=& \int\,\textrm{d}T\,\LDEM_i(T),\label{Ne} \\ 
\label{momento2}
 \Tmi  = \langle \Te \rangle_i &=& \frac{1}{\langle \Ne^2 \rangle_i}\,\int\,\textrm{d}T\,T\,\LDEM_i(T),\label{Tm} 
\end{eqnarray}
}

We define next a measure of the accuracy of the LDEM model to predict the tomographic FBEs in each voxel, as

\begin{equation}\label{R}
R_i \, \equiv \ (1/K) \, \sum_{k=1}^K \left| \, 1 - \FBE^{(k)}_{i,\rm syn}\,/\,\FBE^{(k)}_{i,\rm tom}\, \right| \,,
\end{equation}

being the average relative difference between the tomographic and the synthetic FBEs. The final product of DEMT is in the form of 3D maps of the electron density and mean temperature. These relationships {were} derived in detail in \citealt{frazin_2009} (see Appendix C).\\
%\temp{Aclarar que consideramos Ne como $\sqrt{<\Ne ^2>}$.} {ALBERT: Yo creo que es mejor hablar de $\sqravgN$}

%\temp{DEMT has been improved since the previously published work. We now perform hollow-tomography with the EUV images, (definir hollow-tom) and a 3D regularized of the solution was applyed to obtain smother and more reliable result at the edge of the tomographic grid.} \notebyalbert{NOTA de Albert: Explicá que hacés sin justificarte. Luego aclará como se hizo eso en trabajos previos, y cual es el impacto esperado. Puedo proveer un párrafo de ejemplo, pero creo que será mejor que vos escribas como te salga y yo luego edito. Algo muy importante para todo el paper: proponete escribir ``párrafos modulares". Es decir que, con las razonables limitaciones, puedan moverse de lugar, de orden, y sigan funcionando. No definiría hollow-tomography, es un término coloquial interno nuestro. Lo correcto es que al hablar de datos mucho más arribas expliques que los datos del disco no se utilizan, pero no le des un nombre a la tomografía.}

\subsection{AWSoM simulations}\label{awsom}

\noindent\notebyalbert{This section will summarize the AWSoM model. After a first version is written we will surely require some input from the Michigan friends. Points to cover:
\begin{itemize}
  \item Data input. Link with the data input for DEMT in terms of dates, etc. One paragraph.
  \item Main physical and technical aspects (grid in particular) of the model.  One paragraph. 
  \item Explain the artificially extended TR. One paragraph.
  \item What are the products of the model we care about in this paper. One paragraph.
\end{itemize}
}

In the transition region {(TR)}, the plasma heats up and becomes less dense {by} several orders of magnitude {over} a short distance. Modeling this region {realistically has a very high computational cost}. For this {reason,} the model uses {an artificially} extended {TR about $\approx 0.05\,\mrsun$ thick, as well as artificially large values for the electronic density and temperature as boundary conditions in $r=1\,\mrsun$.} Consequently, model results below {$\approx 1.05\,\mrsun$} were not taken into account in this article.

\subsection{Tracing of results along fieldlines}\label{trace}

\noindent\notebyalbert{This section will explain how the AWSoM magnetic model is used to trace the thermodynamic products of both DEMT and AWSoM along the fieldlines of AWSoM. This requires covering the following points:
\begin{itemize}
  \item A first paragraph qualitatively describing the process. This includes mentioning that the AWSoM products are first interpolated into the DEMT grid.
  \item A short paragraph describing the numerical method used to trace field lines. We have a nice version of this in your previous paper.  At this point the reader should understand that for each field line we have its geometry $\mathbf{r}(s)$ as a function of the distance $s$ along the loop, with $s(1\,\mrsun)\equiv 0$. 
  \item Explain very simply how then that for each voxel the middle point of the field line is selected and assigned the DEMT and AWSoM products of that cell. At this point the reader should understand that for each field line we have: $Q[r(s)]$, where $Q$ is any given quantity from either the DEMT model (specifically $\sqravgN$ and $\Tm$) or the AWSoM model (specifically $\Ne$ and $\Te$).
  \item Explain fits to the DEMT results  $\sqravgN(r)$, $\Tm(r)$.
  \item Explain then the classification of field lines.
\end{itemize}
}

For both rotations, the DEMT and AWSoM 3D maps of electron density $\Ne$, electron mean temperature $\Tm$ and electron temperature $\Te$ respectively were traced along the magnetic field lines modeled using AWSoM as described at the end of Section \ref{demt}. In the case of closed loops, results were separated into their two legs, defined as the two segments that go from the coronal base up to the apex.

For each open field line, and for each leg of the closed ones, an exponential fit was then applied to {the $\sqravgN(r)$ data points, and a linear fit applied to the $\Tm(r)$ data points}, as described by

\begin{eqnarray}\label{Nfit}
\sqravgN &=& N_0\,\exp{\left[\,-\,\left(h/\l\right)\,/\,\left(r/\mrsun\right)\,\right]}\\
\label{Tfit}
\Tm &=& T_0\,+\,m\,h
\end{eqnarray}
\noindent
where $h \equiv r - 1\,\mrsun$ is the coronal height measured from the photosphere. In the electron density fit, $\l\,[\mrsun]$ is the density scale height and $N_0\,[\cminvc]$ is the electron density at the footpoint ($h=0$) of the loop. In the electron temperature fit, $m\,[\MK/\mrsun]$ is the slope and $T_0\,[\MK]$ is the electron temperature at the footpoint of the loop. The slope $m$ is the radial gradient of the electron temperature along the loop, which we denote as $m = \dr\Tm$ hereafter, being $\dr\equiv\er\cdot\nabla$ the radial derivative operator, where $\er$ is the heliocentric radial unit vector.

In the case of the electron density, the fitted function corresponds to the isothermal hydrostatic equilibrium solution, allowing for variation of the solar gravitational acceleration with height. This choice of function provides a straightforward means to directly evaluate how compatible is the observed coronal thermodynamical state with the hydrostatic solution.

%La temperatura en funcion de la altura a lo largo de lineas muestra comportamientos lineales. En el presente trabajo nos focalizamos en lineas con temperaturas crecientes y decrecientes en foma lineal. Es decir nos quedamos con aquellas piernas donde el coeficiente de pearson entre la temperatura y la altura (rho) cumple rho mayor 0.5 (con p-value 0.05) y de esta forma se descartan aquellas piernas en las que los resultados DEMT no pueden asegurar un cambio notable de temperatura entre la base y el apice (asi no uso la palabra isotermico y evado ese problema)

%The temperature as a function of height along legs shows linear behaviors. 
In the present work we focus on lines with increasing and decreasing temperatures in linear form. In other words, we selected those legs where the {Pearson correlation coefficient $\rho$ between temperature $T$ and radius $r$ meets $|\rhoTr| > 0.5$}. \temp{(with p-value less than 0.05)}. In this way, those legs in which the DEMT results cannot ensure a noticeable change in temperature between the base and the apex, are discarded.

The linear fit allows characterization of its variation with height by means of a characteristic temperature gradient $\dr \Tm\,[\MK/\mrsun]$ along each leg.
\diego{En realidad esta regresion lineal utiliza de por si los errores, pero luego tmb los uso para el test de hipotesis, quizas vale la pena ponerlo solo ahi.}

To measure how good the quality of the exponential fit to the electron density and the linear fit to the temperature data are, a chi-square merit function was used (see Section 15.2 \citet{recipes}). The errors were estimated in \citet{lloveras_2017}, $0.07$ MK for $\Tm$ and $4.0\,10^6 cm^{-3}$ for $\Ne$ and legs with p-value $>0.005$ were rejected.

%el p-valor es un concepto estadístico, no veo xq deberia ser explicado. Lo que si, en el paper explicare bien las hipotesis nulas y de esa forma el p-valor queda bien claro.
%El p-valor es una cuantificacion de con cuanta probabilidad de exito estas rechazando o aceptando una hipotesis en particular que se la suele denominar hipotesis nula.
%En este caso, con 95$\%$ de seguridad estas rechazando las piernas que tienen un $pearson_x,y = 0$. Es decir te estarias quedando con aquellas que tienen una dependencia lineal entre x e y.
%En el segundo caso, el test de hipotesis seria sobre la hip nula de que los datos provienen una funcion distribucion donde la recta fiteada de esos datos es el valor esperado de la distribucion. Con lo cual hacerle un test chi cuadrado te da una idea de cuan acertado o no es el ajuste. En este caso la hipotesis nula es que los datos provienen de esa distrib, con lo cual el chi deberia dar bajo y descartas los que dan valores altos con un 95$\%$ de seguridad. Eso es lo que representa el p-valor =0.05 en este caso (el area bajo la funcion distribucion, es decir, probabilidad).

{To be selected a leg must meet all following conditions.}

\begin{enumerate}

\item 
The leg must go through at least five tomographic grid cells with reconstructed data, and there must be at least one data point in each third of the range of heights spanned by the leg.

\item 
The {value of the Pearson correlation coefficient meets $|\rhoTr| > 0.5$.}

\item 
The p-value of both the exponential and linear fit must be less than 0.05.

\end{enumerate}

For the two targeted rotations, DEMT and AWSoM results were traced along the AWSoM field lines, which provides a way to characterize the global thermodynamic state of the inner solar corona and to compare them. To that end four different types of {legs} corresponding with four different types of magnetic structures were selected:
%\temp{Hacer explicacion de los tipos de lineas bien secos, luego en fig 3 hacer notar que implica cada una en cuanto a estructura magnetica}\\

\begin{itemize}
 \item Type 0 - Small (meaning with apex within the DEMT range of heights) closed field lines in the latitudes range $[-30\mdeg,+30\mdeg]$ and $\dr \Tm < 0$. %This class thoroughly populates the equator of the Streamer belt during a minimum solar rotation.

\item  Type I - Small closed field lines in all latitudes, with $\dr \Tm > 0$. %This class thoroughly populates the inner of the Streamer belt.

\item  Type II - Large closed field lines (meaning their apex surpasses the DEMT range of heights) in the latitudes range $[-90\mdeg,-30\mdeg] \cup [+30\mdeg,+90\mdeg]$ and $\dr \Tm > 0$. %This class is mostly populated by very large trans-equatorial field lines forming the envelope of the Streamer belt.

\item Type III - Open field lines (meaning with apex out the DEMT range of heights) in the latitudes range $[-90\mdeg,-60\mdeg] \cup [+60\mdeg,+90\mdeg]$ and $\dr \Tm > 0$.

\end{itemize}
\diego{Hice las definiciones con gradiente de temperatura para que quede mas claro, pero en realidad uso el pearson.} \notebyalbert{Charlemos los tres en persona de esto.}

\subsection{{Energy input flux}}\label{energia}

\noindent\notebyalbert{This section will summarize how loop integrated energy quantities are computed from the DEMT products and the AWSoM magetic model alone. The summary will not include any derivation and will refer to Cecilia's paper. Only the results that are relevant are to be explicited. Start by explaining the model and assumptions, and end up summarizing the loop-integrated products (radiative and conductive fluxes). At the end the reader should understand that their sum is the enery input flux required at the coronal base to maintain the loop stable. This should be a very short section, as it is a summary of one previous paper alone.}

%correcciones a hacer! DIEGO
%Los términos de la Eq (10) requieren ser explicados. La explicación a conituación no es clara, y de hecho está mal escrita. La Eq. (7) se integra en el VOLUMEN CORONAL COMPLETO de cada tubo de flujo y se divide por el área basal del mismo. Utilizando la condición solenoidal de B se llega a la Eq. (7), dónde... y ahí deberían darse las expresiones de Phir y Phic porque no pertenecen a la literatura estandard, y explicar a continuación como todos los elementos de esas dos expresiones surgen de los fits de DEMT y del modelo magnético de AWSoM. Donde si se puede ahorrar mucho espacio es en lo previo, la Eq (8) podría no incluirse por ejemplo. Creo que ahora lo mejor es que trates de completar la explicación dando las expresiones de Phir y Phic, y luego vemos de como limarla, probablemente eliminando la ec. (8). y además el texto puede compactarse, corregirse y reordenarse mucho. Poné las unidades de las cantidades Phi entre paréntesis a continuación de $\Phi_{h,r,c}$.

Due the difference between temperatures of transition region and solar corona, there must to be a heating mechanism that compensate all the losses present in the atmosphere and maintain stable the plasma temperature. Previous works have studied this mechanism by observations and modeling, but mostly for active regions. \citet{maccormack_2017} have developed a technique that estimate the heating injected into the loops during a period of minima activity. Using DEMT combined with an extrapolation of magnetic field, as we detailed before, they obtain the different plasma parameters along each magnetic loop, and therefore, compute the energy input flux required to maintain them stable.
The model assume a simple energy balance for each magnetic flux tube, where the losses by radiative power ($E_r$) and thermal conduction power ($E_c$) are compensated by some coronal heating power($E_h$) \citep{aschwanden_2004}:

\begin{equation}\label{Balance}
E_h(s) = E_r(s)+ E_c(s),
\end{equation}

\noindent
where $s$ is the position along each magnetic flux tube and the powers are in units of $[{\rm erg\,sec^{-1}\,cm^{-3}}]$.

\noindent\notebyalbert{Trabajé todo lo naranja. Primero decidí que es mejor no numerar $E_c$ sino $F_c$. Eso te quedará claro cuando agregues las ecuaciones (11) y (12) que te indico más abajo y leas el último párrafo naranja de abajo, ahí cierra todo. Además describí que son los $\phi$ en forma correcta. El texto previo estaba muy flojón, entre otras cosas se hablaba de una (inexistente) ``conductive loss function" y de una (incorrectamente denominada) ``area along the loop". Obviamente luego de chequear este texto naranja borrá esta nota azul en particular.}

{The thermal conduction power $E_c$ equals the divergence of the conductive heat flux $F_c$, i.e.
$E_c(s)=[1/A(s)]\,{\rm d}[A(s)\,F_c(s)]/{\rm d}s$, where $A(s)$ is the transversal area of the magnetic flux tube at position $s$. Under a quiescent solar corona plasma regime, the conductive flux is assumed to be dominated by the electron thermal conduction, described by the usual Spitzer model \citep{spitzer_1962}

\begin{equation}\label{Fc}
F_c(s)=-\kappa_0\,{T(s)}^{5/2}\,\frac{dT}{ds}(s),
\end{equation}
where $\kappa_0 = 9.2 \times 10^{-7}  {\rm erg\,sec^{-1}\,K^{-7/2}}$ is the Spitzer thermal conductivity.}

Radiative power depends on the amount of plasma in certain temperatures that can radiate. The model estimate it by integrating the squared density multiplied by a radiative loss function $\Lambda(T)$. This function depends on plasma temperature and is calculated by the atomic database and the plasma emission model from CHIANTI \citep{delzanna_2015}. Then, the expression for radiative power obtained is:

\begin{equation}\label{Er}
E_r = \int\,\textrm{d}T\,\LDEM(T)\,\Lambda(T)
\end{equation}

{The energy balance given by Equation (\ref{Balance}) is then integrated in the volume of any given coronal magnetic flux tube. Dividing the result by the flux tube area at the coronal base, and making use of the soleidonal conditon of the magnetic field, a field-line integrated version of that energy balance is found,
\begin{equation}\label{FluxBalance}
\phi_h = \phi_r + \phi_c.
\end{equation}
\noindent
where the line integrated flux quantities $\phi_{r,c}\,[\rm{erg\,sec^{-1}\,cm^{-2}}]$ are given by \citep{maccormack_2017},} \\

\begin{equation}\label{phi_r}
\phi_r = \left( \frac{B_0 \ B_L}{B_0 + B_L} \right) \int_{0}^{L} ds \ \frac{E_r(s)}{B(s)}
\end{equation}

\begin{equation}\label{phi_c}
\phi_c = \left( \frac{B_0 \ F_{c,L} - B_L \ F_{c,0}}{B_0 + B_L} \right) 
\end{equation}

{Note that, for any given field line, all quantities in these two expresions can be computed from the DEMT and AWSoM models through Equations (\ref{Fc})-(\ref{Er}). Once computed, the quantity $\phi_h$ can be calculated, which is the energy input flux required at the coronal base of each coronal field-line to mantain a stable coronal structure.}

\section{Results}\label{resu}

\subsection{Tomographic Results}\label{demt_res}

\noindent\notebyalbert{This section will show first the DEMT products in the form of Carrington maps at three heights, for the two target rotations. It will then show the latitude/longitude location of the four types of  magnetic loops (defined in Section 2.3) at a sample height. Then, for the each type of loop, we will show histograms of the values of different DEMT products traced along them, charerizing in a quantitative fashion the DEMT results in distinct magnetic structures. It will also include a quantitative comparison between the two target rotations of the median values of the DEMT products in each type of loop.}

{To perform a comparative study of the thermodynamic structure of the inner corona in different solar cycles
we carry out a DEMT reconstruction for two rotations: CR-2082, using STEREO/EUVI data and CR-2208, using SDO/AIA data.} {Once the LDEM was determined for each rotation, the square root of the mean value of the electron density squared $\sqravgN$ and the electron mean temperature $\Tm$ were computed at each voxel of the tomographic computational grid by means of Equations \ref{momento1} and \ref{momento2}.}

\albert{As an example,} Figures \ref{carmaps_demt_2082} and \ref{carmaps_demt_2208} show \albert{latitude-longitude} maps of DEMT results for both \albert{rotations. Three different heights of interest were selected here, providing also a detailed 3D view of the tomographic results: the lowest height of tomographic results ($1.025\,\mrsun$), one of the lowest heights where the AWSoM results are consistent with coronal conditions ($1.065\,\mrsun$), and a middle height of the tomographic grid ($1.105\mrsun$)}. Black voxels correspond to non-reconstructed voxels (\albert{see} Section \ref{demt}). Thick-black \albert{curves indicate} the magnetic open/closed \albert{boundaries} based on the magnetic field of \albert{the} AWSoM model.

\albert{Target} CR-2082 was highly axisymmetric \albert{(i.e, a high azymuthal symmetry)} and showed two ARs (active region) in latitude $+30\mdeg$ and longitude $50\mdeg$ and $120\mdeg$. \albert{Target CR-2208 was much less axisymmetric and it} showed two AR, one around latitude $+5\mdeg$ and longitude $140\mdeg$ and the other around latitude-longitud $+5\mdeg$, $300\mdeg$. \albert{In both rotations the open/closed} boundaries derived from the respective AWSoM model \albert{very accurately} match the shape of \albert{contour levels} of the tomographic maps of both the \albert{electron density and temperature. The magneticlly open regions of the AWSoM model are those associated to CHs, and the magnetically closed ones to the equatorial streamer belt.}

\albert{Visual inspection of Figures \ref{momento1} and \ref{momento2} indicates} that the streamer belt region of CR-2082 was denser and colder than CR-2208. \albert{In the case of CR-2082, which belongs to the deep minimum epoch between SC2 24/25, the temperature in the equatorial latitudes tends to decrease with height while in the mid-latitudes tends to increase. A similar behabiour is seen in CR-2208, belonmging to the declining phase of SC25, but being a much less axisymmetric target this behaviour is not so obvious. This structure of the streamer have been reported for other solar minimum rotations in previous DEMT works \citep{lloveras_2017,nuevo_2013,vasquez_2010}.} Closed regions are characterized by relatively larger values of density and temperature in comparison to the open regions.

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.495\textwidth]{figs/map_Ne_CR2082_DEMT-EUVI_behind_H1-L3523_r3d_1025_Rsun.pdf}
\includegraphics[width=0.495\textwidth]{figs/map_Tm_CR2082_DEMT-EUVI_behind_H1-L3523_r3d_1025_Rsun.pdf}
\includegraphics[width=0.495\textwidth]{figs/map_Ne_CR2082_DEMT-EUVI_behind_H1-L3523_r3d_1065_Rsun.pdf}
\includegraphics[width=0.495\textwidth]{figs/map_Tm_CR2082_DEMT-EUVI_behind_H1-L3523_r3d_1065_Rsun.pdf}
\includegraphics[width=0.495\textwidth]{figs/map_Ne_CR2082_DEMT-EUVI_behind_H1-L3523_r3d_1105_Rsun.pdf}
\includegraphics[width=0.495\textwidth]{figs/map_Tm_CR2082_DEMT-EUVI_behind_H1-L3523_r3d_1105_Rsun.pdf}
\caption{Carrington maps of DEMT {products $\sqravgN$} (left panels) and $\Tm$ (right panels) for CR-2082. Top, middle and bottom panels show the results at three heliocentric heights, $1.025$, $1.065$ and $1.105\,\mrsun$ respectively. Black voxels correspond to non-reconstructed regions (see text in Section \ref{demt_res}) and thick-black curves indicate the open/closed boundaries.}
\label{carmaps_demt_2082}
\end{center}
\end{figure}

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.495\textwidth]{figs/map_Ne_CR2208_DEMT-AIA_H1_L522_r3d_1025_Rsun.pdf}
\includegraphics[width=0.495\textwidth]{figs/map_Tm_CR2208_DEMT-AIA_H1_L522_r3d_1025_Rsun.pdf}
\includegraphics[width=0.495\textwidth]{figs/map_Ne_CR2208_DEMT-AIA_H1_L522_r3d_1065_Rsun.pdf}
\includegraphics[width=0.495\textwidth]{figs/map_Tm_CR2208_DEMT-AIA_H1_L522_r3d_1065_Rsun.pdf}
\includegraphics[width=0.495\textwidth]{figs/map_Ne_CR2208_DEMT-AIA_H1_L522_r3d_1105_Rsun.pdf}
\includegraphics[width=0.495\textwidth]{figs/map_Tm_CR2208_DEMT-AIA_H1_L522_r3d_1105_Rsun.pdf}
\caption{Same as Figure \ref{carmaps_demt_2082} but for CR-2208.}
\label{carmaps_demt_2208}
\end{center}
\end{figure}

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.495\textwidth]{figs/map_R_CR2082_DEMT-EUVI_behind_H1-L3523_r3d_1025_Rsun.pdf}
\includegraphics[width=0.495\textwidth]{figs/map_R_CR2208_DEMT-AIA_H1_L522_r3d_1025_Rsun.pdf}
\includegraphics[width=0.495\textwidth]{figs/map_R_CR2082_DEMT-EUVI_behind_H1-L3523_r3d_1065_Rsun.pdf}
\includegraphics[width=0.495\textwidth]{figs/map_R_CR2208_DEMT-AIA_H1_L522_r3d_1065_Rsun.pdf}
\includegraphics[width=0.495\textwidth]{figs/map_R_CR2082_DEMT-EUVI_behind_H1-L3523_r3d_1105_Rsun.pdf}
\includegraphics[width=0.495\textwidth]{figs/map_R_CR2208_DEMT-AIA_H1_L522_r3d_1105_Rsun.pdf}
\caption{{Carrington maps of the measure $R$ defined by Equation (\ref{R}), for CR-2082 (left panels) and CR-2208 (right panels), at heights $1.025$, $1.065$ and $1.105\,\mrsun$, from top to bottom}. }
\label{carmaps_R_2082_2208}
\end{center}
\end{figure}

{To study the variation of the DEMT results in different magnetic structures we traced $\sqravgN$ and $\Tm$ along the magnetic field lines of the AWSoM model.} For both rotations, all field lines that meet the criteria listed in Section \ref{trace} were selected. For each field line, the data points of electron density and electron mean temperature as a function of height were fitted to the Equations \ref{Nfit} and \ref{Tfit}. \albert{As a result, the electron density $N_0(r=1.0\,\mrsun)$ and scale height $\lN$ were computed for each leg, as well as the temperature gradient $\dr\Tm$, and the height-averaged (along the leg) {electron temperature} $\aTm$.}

\noindent\notebyalbert{Quiero destactar el siguiente ejemplo de corrección, todo el párrafo. Comparalo con el texto previo para ver cuanto cambió. No solo estaba mal escrito sintácticamente, sino que no se conectaba adecuadamente con cada aspecto puntual de la Sección 2.3. Fijate que quedó mucho mejor explicado y conectando bien puntualmete con elementos previos de tu paper, lo cual es importante. Quizá te sirve para mejorar párrafos puntuales. Moví al fin de este párrafo el detalle de los porcentajes de cada población.}

\albert{For both target rotations, the top panels of Figure \ref{rpoint_demt} show the latitude-longitude location (at heliocentric height $r=1.105\,\mrsun$) of all traced field line legs for which criterion (i) of Section \ref{trace} is met. Open legs are indicated in grey color and closed ones in black color. For each leg, the fits to tomografic temperature and density were applied, as given by Equations (\ref{Nfit})-(\ref{Tfit}). Considering the DEMT data points and the resulting fits along each leg, the bottom panels of Figure \ref{rpoint_demt} show the latitude-longitude location of the subset for which also both criteria (ii) and (iii) of Section \ref{trace} are met. Using a four-color code, type 0, I, II and III legs are shown in blue, red, violet and green color, respectively.} \albert{Of the 45000 legs selected for CR-2082, 14\% are type 0, 38\% are type I, 23\% are type II and 25\% type III. On the other hand, of the 59000 legs selected for CR-2208, 8\% are type 0, 43\% are type I, 22\% are type II and 30\% type III.}

\albert{Type 0 (small down) legs mainly populate the equatorial latitudes. This kind of structure was originally found by \citet{huang_2012}, and their existence was shown to be anti-correlated with the solar activity level around the SC 24/25 solar minimum by \citet{nuevo_2013}. Later on, \citet{lloveras_2017} showed how equatorial down loops in streamers were also to be found in the deep minimum between SCs 23/24. Here, we verify the existence of this type of structuire for the two target rotations. The relatively smaller population of down loops seen in CR-2208, as compared to CR-2082, is consistent with the aforementioned results by \citet{nuevo_2013}. Type I (small up) mainly populate the mid-latitudes, while type II (large up) legs are mostly very large trans-equatorial field lines forming the envelope of the streamer belt. Finally, type III (open) legs populate of course the CHs.}

\notebyalbert{No comprendo esta frase:} \emph{This classification of types of lines automatically leads to specific regions arranged along all longitudes associated with different thermodynamic, with good symmetry between both hemispheres in both rotations.}

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.495\textwidth,clip=]{figs/Grispoint_2082_demt_paper_test_Rpoint-map.pdf}
\includegraphics[width=0.495\textwidth,clip=]{figs/Grispoint_2208_demt_paper_test_Rpoint-map.pdf}
\includegraphics[width=0.495\textwidth,clip=]{figs/Highpoint_2082_demt_paper_cr2082_full_Rpoint-map.pdf}
%\includegraphics[width=0.495\textwidth,clip=]{figs/Highpoint_2082_demt_paper_cr2082_up_Rpoint-map.eps}
%\includegraphics[width=0.495\textwidth,clip=]{Highpoint_2082_awsom_paper_Rpoint-map.eps}
\includegraphics[width=0.495\textwidth,clip=]{figs/Highpoint_2208_demt_paper_cr2208_full_Rpoint-map.pdf}
%\includegraphics[width=0.495\textwidth,clip=]{figs/Highpoint_2208_demt_paper_cr2208_up_Rpoint-map.eps}
%\includegraphics[width=0.495\textwidth,clip=]{Highpoint_2208_awsom_paper_Rpoint-map.eps}
\caption{\emph{Top panels:} \albert{latitude-longitude location at heliocentric height $r=1.105\,\mrsun$ of all open (grey color) and closed (black color) traced field line legs for which criterion (i) of Section \ref{trace} is met, for both CR2082 (left) and CR2208 (right). \emph{Bottom panels}: latitude-longitude location of the subset for which also both criteria (ii) and (iii) of Section \ref{trace} are met. Using a four-color code, type 0, I, II and III legs (see text) are shown in blue, red, violet and green color, respectively.}}
\label{rpoint_demt}
\end{center}
\end{figure}

\albert{Figure \ref{gradt_demt} shows frequency histograms (normalized to the total amount of selected field lines ) of the temperature radial gradient ($\dr\Tm$) for legs of type 0, I, II and III.} The lack of population around values close to zero is due to the requirement $|\rho_t| > 0.5$ which discards quasi-isothermal legs. \albert{For both rotations, the median value of $\dr\Tm$ is very similar for each type of leg, being  $\approx -2.5$, $2.3$ and $2.4\,\MK/\mrsun$ for legs of type 0, I and II, respectively. The only notable difference between both rotations is the characteristic value $\dr\Tm\approx 4.5\,\MK/\mrsun$ for legs of type III for CR-2208, which is about half that for CR-2082.} \notebyalbert{Estos vaores que informas son medianas o modas? (yo especifiqué ``medianas". Son efectivamente tan iguales?}

%\temp{mostrar los gradientes negativos va a ser util en los parrafos posteriores que expliquen porque en las lineas tipo 0 hay valores de $\phi_c$ negativos.}

%\temp{ Comentar en la discusion el paper de shift y cramer, para decir un poco sobre los down y la incapacidad del modelo en reproducirlos}

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.495\textwidth,height=0.25\textwidth,clip=]{figs/histo_cr2082_fulltriple_gradt.eps}
\includegraphics[width=0.495\textwidth,height=0.25\textwidth,clip=]{figs/histo_cr2208_fulltriple_gradt.eps}
\caption{Frequency histograms of the temperature radial gradient for the different {four types of legs} shown in Figure \ref{rpoint_demt} (using the same color code) for CR-2082 (left panel) and CR-2208 (right panel).}
\label{gradt_demt}
\end{center}
\end{figure}

\albert{Figure \ref{histos_fulldemt} shows} the statistical results of both rotations for all field lines types. From top to bottom \albert{results are shown for} type 0 to type III \albert{field lines}. From left to right the panels show the statistical distribution of $\NCB \equiv \sqravgN(r=1.055)$ \albert{(the lowest height where the AWSoM results are consistent with coronal conditions)}, $\lN$ \albert{and} $\aTm$, \albert{with the median value $m$ indicated in each plot}.

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.31\textwidth,height=0.18\textwidth,clip=]{figs/histo_2082_2208_fulldemt_streamer_down_ne_1055.eps}
\includegraphics[width=0.31\textwidth,height=0.18\textwidth,clip=]{figs/histo_2082_2208_fulldemt_streamer_down_lambda_n.eps}
\includegraphics[width=0.31\textwidth,height=0.18\textwidth,clip=]{figs/histo_2082_2208_fulldemt_streamer_down_Tm.eps}
\includegraphics[width=0.31\textwidth,height=0.18\textwidth,clip=]{figs/histo_2082_2208_fulldemt_streamer_up_ne_1055.eps}
\includegraphics[width=0.31\textwidth,height=0.18\textwidth,clip=]{figs/histo_2082_2208_fulldemt_streamer_up_lambda_n.eps}
\includegraphics[width=0.31\textwidth,height=0.18\textwidth,clip=]{figs/histo_2082_2208_fulldemt_streamer_up_Tm.eps}
\includegraphics[width=0.31\textwidth,height=0.18\textwidth,clip=]{figs/histo_2082_2208_fulldemt_bound_up_ne_1055.eps}
\includegraphics[width=0.31\textwidth,height=0.18\textwidth,clip=]{figs/histo_2082_2208_fulldemt_bound_up_lambda_n.eps}
\includegraphics[width=0.31\textwidth,height=0.18\textwidth,clip=]{figs/histo_2082_2208_fulldemt_bound_up_Tm.eps}
\includegraphics[width=0.31\textwidth,height=0.18\textwidth,clip=]{figs/histo_2082_2208_fulldemt_CH_up_ne_1055.eps}
\includegraphics[width=0.31\textwidth,height=0.18\textwidth,clip=]{figs/histo_2082_2208_fulldemt_CH_up_lambda_n.eps}
\includegraphics[width=0.31\textwidth,height=0.18\textwidth,clip=]{figs/histo_2082_2208_fulldemt_CH_up_Tm.eps}
\caption{{Statistical distribution of DEMT results traced along legs of type 0, I, II and III (from top to bottom), as defined in Section \ref{trace}. From left to right: electron density at the coronal base $\NCB\equiv\sqravgN(r=1.055\,\mrsun)$, electron density scale height $\lN$, and loop-averaged electron temperature $\aTm$. Results for target rotations CR2082 and CR2208 are shown in blue and red color, respectively. In each panel the median value $m$ is indicated.} \notebyalbert{ (by Albert) The histogram Figures in this draft need to be improved, in particular font size-and-type need to be optimized for clear reading. Also, we need to improve their X/Y ratio, here I stretched the Y size a bit in the LaTeX source.} }
\label{histos_fulldemt}
\end{center}
\end{figure}

For all type of lines, Table \ref{tabla_demt} \albert{summarizes a} quantitative comparative analysis between the results of the two \albert{target rotations}. For CR-2082 quantities are expressed as absoulte values, \albert{while} for CR-2208 they are informed \albert{as a precentual variation} relative to the corresponding results for CR-2082. The following major results concerning the structure of each rotation individually and their comparison can be drawn.

\begin{table}
\begin{tabular}{l r@{.}l@{\hskip 0.05in} r@{\hskip 0.01in} r@{.}l  r@{.}l@{\hskip 0.05in} r@{\hskip 0.01in} r@{.}l r@{.}l@{\hskip 0.05in} r@{\hskip 0.01in} r@{.}l }
% r@{.}l@{\hskip 0.05in} r@{\hskip 0.01in} r@{.}l@{\hskip 0.01in} l}
\hline
Type    & \multicolumn{5}{c}{$\med(\NCB)$}             & \multicolumn{5}{c}{$\med(\lN)$} & \multicolumn{5}{c}{$\med(\avgTe)$} \\
       & \multicolumn{5}{c}{$[10^8\,{\rm cm}^{-3}]$}  & \multicolumn{5}{c}{$[{\rm 10}^{-2}\,\mrsun]$} & \multicolumn{5}{c}{$[\MK]$} \\
\hline
0    & 1&26 &(\Mi&17&4\%)  &   7&3 &(\Pl&13&6\%) &   1&09 &(\Pl&24&7\%) \\
I    & 1&13 &(\Mi&14&1\%)  &   8&3 &(\Pl&18&1\%) &   1&29 &(\Pl&20&1\%) \\
II   & 0&97 &(\Mi&20&6\%)  &   8&9 &(\Pl&29&1\%) &   1&33 &(\Pl&16&5\%) \\
III  & 0&66 &(\Mi&16&2\%)  &   7&7 &(\Pl&11&4\%) &   0&98 &(\Pl&~6&1\%) \\
\hline          
\end{tabular}
\caption{Median value (indicated as ``Md'') of the statistical distribution of $\NCB$, $\lN$, and $\aTm$ for each coronal type of lines defined in Section \ref{trace}. For CR-2082 values are expressed in absolute terms, while for CR-2208 they are informed as a percentual variation relative to the CR-2082 value. \albert{Propongo que para todas las cantidades absolutas informes siempre dos cifras significativas, y que todos los porcentajes los redondees a cero decimales (o sea, a veces una y a veces dos cifras significativas) y que no uses puntos para estos nunca. Doy dos ejemplos para no dejar dudas. Para $\NCB$ de Type-II: $1.0\,(-21\%)$. Para $\aTm$ de Type-III: $1.0\,(+6\%)$.}}
\label{tabla_demt}
\end{table}

%\noindent\diego{Si queda una asociacion directa entre los tipos de lineas y las estructuras de streamer, CH etc entonces usaria esos nombres a continuacion.}\\
\albert{Throughout} the equatorial streamer of both rotations, \albert{type 0, I and II legs, which are associated to progressively} increasing latitudes, \albert{progressively exhibit} decreasing {coronal base} density, \albert{increasing density scale height, and increasing electron temperature}. In both rotations the CH regions are characterized sub-MK temperatures, and coronal base densities of order $\approx 1/2$ of \albert{those observed for the type 0 lines in the core of} the equatorial streamer. \notebyalbert{No estoy seguro de entender y acordar con la siguiente frase.} \emph{Histograms with sharp populations indicates that for both rotations the values of all physical quantities exhibit a strong symmetry between both hemispheres.}

\notebyalbert{A continuación modifiqué la sintaxis en que esto se explica (creo que para mejor), y puse números acordes a los de la Tabla-1, los que estaban no eran consistentes con ella.} \albert{A comparison of the results between the two rotations shows that compared to CR-2082, target rotation CR-2208 was characterized by $\approx 15-20\%$ lower values of the electron density at the coronal base, $\approx 10-30\%$ larger values of density scale height, and $\approx 15-25\%$ larger values of the electron temperature.}

%To establish if the differences in the results between the two rotations are significant, Section \ref{uncertainties} estimates the error bars due the dominating systematic uncertainties.
%\diego{quizas vale la pena introducir alguna comparacion sobre el campo magnetico, veremos como quedan los comentarios de ceci.}
%{Also, within the streamer belt the magnetic field strength is similar in both rotations in the northern hemisphere, while in the southern hemisphere CR-1915 exhibits $10-80\%$ larger values compared to CR-2081, with increasing difference for larger latitudes. In the coronal holes, the magnetic field strength of CR-1915 is much larger than for CR-2081, exhibiting values up to 50\% larger in the northern hemisphere and more than twice larger in the southern hemisphere.}

%Previous works have found magnetic loops which temperature decrease with height, most generally in equatorial latitudes \citep{huang_2012,nuevo_2013}. This type of loops with negative gradients produce a negative conductive flux that contribute to the heating in the corona. In this work, we compute the loop-integrated energy quantities separately in three different populations of loops.

\notebyalbert{Cambié este párrafo.} \albert{To analyze the loop-integrated energy flux quantities, we selected closed loops where both legs meet the requirements of Section \ref{trace}, and both posses the same sign of the radial gradient of the electron temperature $\dr\Tm$. In this way, according to the clasification of both its legs, each given loop was classified as of type 0 (small down loop), I (small up loop), or II (large up loop).}

\begin{figure}[ht]
\begin{center}
\includegraphics[width=0.495\textwidth,clip=]{figs/histocr2082_ccdownenergia.eps}
\includegraphics[width=0.495\textwidth,clip=]{figs/histocr2208_ccdownenergia.eps}
\includegraphics[width=0.495\textwidth,clip=]{figs/histocr2082_ccenergia.eps}
\includegraphics[width=0.495\textwidth,clip=]{figs/histocr2208_ccenergia.eps}
\includegraphics[width=0.495\textwidth,clip=]{figs/histocr2082_cgenergia.eps}
\includegraphics[width=0.495\textwidth,clip=]{figs/histocr2208_cgenergia.eps}
\caption{\albert{Sugiero fuertemente que el histograpma de Type 0 tenga xr=[-1,+3] como el resto, creo que quedará claro por qué cuando lean el análisis que hice de esta Figura en el texto.} {Statistical results of the loop-integrated energy flux quantities $\phi_r,\phi_c,$ and $\phi_h$ in colors blue, red and green, respectively for CR-2082 (left) and CR-2208 (right). From top to bottom, panels show the results for loops of type 0, I and II, which are loops for which both legs meet the criteria from Section \ref{trace}.} \notebyalbert{(by Albert) En el XYOUTS podés poner la letra griega $\phi$ usando los comandos ``!4" y ``!3",  y para que te salga r,c,h como subíndice pero no tan pequeño usá: ``!Dp!N", dónde p=r,c,h.}}
\label{energia_demt}
\end{center}
\end{figure}

For both target rotations, and for loops of type 0, I and II, Figure \ref{energia_demt} shows the frequency histogram of the loop-integrated energy flux quantities \albert{$\phi_r$, $\phi_c$ and $\phi_h$ in blue, green and red color}, respectively. \notebyalbert{Notá que puse las cosas en el orden que las cosas se calculan: Eq. 11, 12, y 10. Sugiero que en los histogramas listes los $\phi$ y sus medianas de arriba a abajo en este orden, ya que igual los deberás re-hacer para poner la letra griega $\phi$ (y otros cambios requeridos).}

\notebyalbert{El análisis que había de los $\phi$ no me pareció adecuado, y lo comenté (no lo borré por si querés revisarlo y si ves que se pasó algo). Lo que sigue hasta el fin de la sección lo escribí yo mirando el paper de Cecilia y tu Figura, destacando lo que considero más importante. Lo hice medio apurato, y requerirá más trabajo. Pero para mi esta es la línea.}

\albert{For both rotations it is readily seen that the integrated radiative loss of the loops, measured by the quantity $\phi_r$, is largest for loops of type 0, smaller for loops of type I, and smalest for loops of type II . This is mainly due to the fact that in the range of sensitivity of the AUVI and AIA instruments, namely 0.5–3.0 MK \citep{nuevo_2015}, the radiative loss function $\Lambda(T)$ has a local maximum at $\Tc\approx 1\,\MK$. Indeed, according to Table 1, loops of type 0 have the lowest temperature of them all and closest to $\Tc$, for both rotations.}

\albert{In the case of the quantity $\phi_c$, its sign depends on that of the conductive flux $F_c$. As shown in \citet{maccormack_2017}, down loops (type 0) and up loops (type I and II) are characterized by $\phi_c<0$ and $\phi_c>0$, respectively, which is verified in Figure \ref{energia_demt}.}

\albert{The characteristic energy input flux at the coronal base is in the range $\phi_h\approx 0.5-1.0 \times 10^5\,\erg\cminvs\s^{-1}$, depending on the rotatin and the type of loop. These values are consistent with those reported by \citet{maccormack_2017}.}

%Type 0 loops present slightly larger gradients than type I. Therefore, type 0 present bigger amount, in modulus, of loop-integrated conductive flux. Due conductive flux negative can not be assumed as a loss but it can be contributing to the coronal heating. Thus, to compensate the radiative losses, a lower loop-integrated energy input flux ($\phi_h$) is required.

%We can see, on type 0 loops, a median value of energy input flux for CR2082 twice bigger than CR2208. Also, in both rotations, we observe a negative population of loop integrated energy input flux. This can refer to some loops where loop-integrated radiative flux is too small in comparison with the larger gradients of negative conductive flux, resulting in negative energy input flux.

%The type I loops show a loop-integrated positive conductive for each loop in both rotations consistent with positive gradients of temperatures. The conductive flux have values around 0.3 $10^{5}[\rm{erg\,sec^{-1}\,cm^{-2}}]$ in both rotations. Similar results have been found for loop-integrated radiative flux and energy input flux for both rotations.

\subsection{Tomography and Models}\label{awsom_res}

\noindent\notebyalbert{This section will show the results of AWSoM for $N_e$ and $T_e$, and how they compare with the corresponding results of DEMT ($\sqravgN$ and $\Tm$). It will first show the Carrington maps of $N_e$ and $T_e$ at the same three heights we showed the same maps for the DEMT results. Then it will show the lat/lon location of loops of type I, II and III at a middle height, just loke for DEMT results. As the two targets are highly axis-symmetric, a first quantitative comparison between AWSoM and DEMT will be shown as longitude-averaged latitudinal profiles of the results of both models at a sample height. To make comparisons in distinct magnetic structures, we will also show the average radial dependence of results for loops of type I, II and III, overplotting the AWSoM and the DEMT results. For the tree types of loops we will show histograms of the DEMT and AWSoM products along field lines. This is a LOT of information from which and we need to pull out a few clear conlcusions.}

The thermodynamic results of AWSoM are shown in Figures \ref{carmaps_awsom_2082} and \ref{carmaps_awsom_2208}. That is latitude-longitude maps for $\Ne$ and $\Te$ modeled for CR-2082 and CR-2208 at $r=1.025, 1.065$ and $1.105\,\mrsun$. The black curve denotes the same open-close boundary showed in Figures \ref{carmaps_demt_2082} and \ref{carmaps_demt_2208} respectively.

%$r=1.025, 1.065$ and $1.105\,\mrsun$. The black curve denotes the open-close boundary.

A visual inspection of these maps shows that both reconstructions are very axysimmetric for $\Ne$ and $\Te$ as well. The closed region is denser and hotter than the open region. The density and temperature gradients are maximum in the open-close boundary. The open field regions in low latitudes maintain thermodynamic values associated with open regions. The reconstruction of CR-2082 does not show clear the ARs, while CR-2208 shows a well reconstruction of both AR around latitude-longitud $+5\mdeg$, $140\mdeg$ and $+5\mdeg$, $300\mdeg$. In both rotations the open-closed boundaries derived from the AWSoM model match the shape of the contour levels of both the electron density and the electron temperature.

%\diego{comparacion visual entre estas dos figuras y als dos de DEMT.}

The same intensity scales in both density and temperature can be observed between the tomographic results and those of AWSoM in both rotations, as well as the same thermodinamic structures. That is, equatorial latitudes, mid laditudes and the open regions.

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.495\textwidth]{figs/map_Ne_awsom_2082_185_short_1025_Rsun.pdf}
\includegraphics[width=0.495\textwidth]{figs/map_Te_awsom_2082_185_short_1025_Rsun.pdf}
\includegraphics[width=0.495\textwidth]{figs/map_Ne_awsom_2082_185_short_1065_Rsun.pdf}
\includegraphics[width=0.495\textwidth]{figs/map_Te_awsom_2082_185_short_1065_Rsun.pdf}
\includegraphics[width=0.495\textwidth]{figs/map_Ne_awsom_2082_185_short_1105_Rsun.pdf}
\includegraphics[width=0.495\textwidth]{figs/map_Te_awsom_2082_185_short_1105_Rsun.pdf}
\caption{Carrington maps of density (left panels) and temperature (right panels) obtained with AWSoM model for the same three heights shown in Figure \ref{carmaps_demt_2082}.}
\label{carmaps_awsom_2082}
\end{center}
\end{figure}

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.495\textwidth]{figs/map_Ne_awsom_2208_185_short_1025_Rsun.pdf}
\includegraphics[width=0.495\textwidth]{figs/map_Te_awsom_2208_185_short_1025_Rsun.pdf}
\includegraphics[width=0.495\textwidth]{figs/map_Ne_awsom_2208_185_short_1065_Rsun.pdf}
\includegraphics[width=0.495\textwidth]{figs/map_Te_awsom_2208_185_short_1065_Rsun.pdf}
\includegraphics[width=0.495\textwidth]{figs/map_Ne_awsom_2208_185_short_1105_Rsun.pdf}
\includegraphics[width=0.495\textwidth]{figs/map_Te_awsom_2208_185_short_1105_Rsun.pdf}
\caption{Same as Figure \ref{carmaps_awsom_2208} for CR-2208.}
\label{carmaps_awsom_2208}
\end{center}
\end{figure}

%\temp{Se presentan los midpoints para awsom que lucen mas rellenos xq no hay down loops y todo es super suave.}

In the same way as in Section \ref{demt_res}, the model results were traced along the magnetic lines obtained with it. This allowed to obtain $\Ne(r)$ and $T_e(r)$ along the magnetic lines and fit them to the Equations \ref{Nfit} and \ref{Tfit}.

The actual version of the AWSoM model can not reproduce down legs, so for a further comparison we only select lines of type I, II and III of those meeting the criteria described in Section \ref{trace}.

As a way to understand the distribution of legs, Figure \ref{rpoint_awsom} shows the latitudinal-longitudinal location of the traced legs at top and the selected fitted legs meeting the criteria at $1.105\,\mrsun$ for both rotations, in the same way as in Section \ref{demt_res}. In red the type I lines, populating mostly the entire streamer. In violet the type II lines, populating the edge that assemble the envelope around the streamer, and in green dots the type III lines, populating the coronal holes.

The reader may notice that these structures are the same as those in Figure \ref{rpoint_demt} but more filled. The cause is twofold, on the one hand the same magnetic model was used to trace both DEMT and AWSoM results, on the other hand is because giving the model smoother results, fewer fitted lines are filtered.

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.495\textwidth,clip=]{figs/Grispoint_2082_awsom_paper_test_Rpoint-map.pdf}
\includegraphics[width=0.495\textwidth,clip=]{figs/Grispoint_2208_awsom_paper_test_Rpoint-map.pdf}
%\includegraphics[width=0.495\textwidth,clip=]{Highpoint_2082_demt_paper_Rpoint-map.eps}
\includegraphics[width=0.495\textwidth,clip=]{figs/Highpoint_2082_awsom_paper_cr2082_up_Rpoint-map.pdf}
%\includegraphics[width=0.495\textwidth,clip=]{Highpoint_2208_demt_paper_Rpoint-map.eps}
\includegraphics[width=0.495\textwidth,clip=]{figs/Highpoint_2208_awsom_paper_cr2208_up_Rpoint-map.pdf}
\caption{Same as Figure \ref{rpoint_demt}, but using the density and temperature of the AWSoM model to classify {its legs in types I, II and III. The model does not exhibit legs of type 0.}}
\label{rpoint_awsom}
\end{center}
\end{figure}

To highlight the latitudinal structures observed in Figures \ref{carmaps_demt_2082} and \ref{carmaps_awsom_2082} for CR-2082 as well as Figures \ref{carmaps_demt_2208} and \ref{carmaps_awsom_2208} for CR-2208, Figure \ref{perf_lat} shows for both rotations the latitudinal variation  of both the electron density and the electron mean temperature at the height $r=1.105\,\mrsun$, averaged over all longitudes (excluding the AR). DEMT results are shown in blue lines while AWSoM results in red. While the temperature shows a maximum in each hemisphere, the density has a simpler behavior, with higher values in the low-latitudes of the streamer and a nearly monotonic decrease towards larger latitudes. The vertical black lines denotes the average of the open-close boundary in both hemispheres to each rotation.

\diego{Me gustaria hacer un comentario de cierre para esta figura que refiera las Figuras 3 y 9, como la linea de abajo, y que conecte con los histogramas que vienen a continuacion. Si no queda bien, hay que mover esta figura entre la actual figura 8 y 9.}

This longitude-averaged behaviors match well with the physical locations of the legs shown in Figures \ref{rpoint_awsom} and \ref{rpoint_demt}.

According to the visual inspection on the maps of carrington, the latitudinal profiles comparing both DEMT and AWSoM results are on the same scale. Little difference in density is observed in CR-2208 and in temperature in CR-2082, while DEMT in CR-2082 was $20\%$ denser. Finally, DEMT results for CR-2208 were $25\%$ hotter in the streamer and $30\%$ colder in the polar zone than AWSoM.
Density profiles on AWSoM show an increase around the polar zone, which seems like an artifice.

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.495\textwidth]{figs/Perfil_Ne_demt_awsom_2082_1105.eps}
\includegraphics[width=0.495\textwidth]{figs/Perfil_Ne_demt_awsom_2208_1105.eps}
\includegraphics[width=0.495\textwidth]{figs/Perfil_Te_demt_awsom_2082_1105.eps}
\includegraphics[width=0.495\textwidth]{figs/Perfil_Te_demt_awsom_2208_1105.eps}
\caption{Longitude-averaged latitudinal dependence of the electron density (top) and temperature (bottom) for DEMT (blue) and AWSoM (red) results at $1.105\,\mrsun$. The left (right) panels correspond to CR-2082 (CR2208). The vertical black line indicates the longitude-averaged latitudes of the open/close magnetic boundary in both hemispheres. \notebyalbert{(by Albert): The idea of this Figure is to show how DEMT and AWSoM compare AS A FUNCTION OF LATITUDE at an intermediate height of the DMET grid, averaged in longitude (excluding zones with ARs).}}
\label{perf_lat}
\end{center}
\end{figure}

To summarize the comparison between both results during both rotations at full height, it is shown in Figure \ref{perfiles_promedio} the average fits of $\Ne(r)$ and $T_e(r)$ for legs of type I (blue), II (red), and III (green) for both rotations.

In average all temperature and density behavior from both results are on the same scale.
The coronal base densities of AWSoM were lower than those of DEMT in CR-2082 and higher in CR-2208. The fact that the density decays to similar values between both results suggests a similar height scale. Onthe other hand basal temperature of AWSoM was lower in both rotations, but the temperature gradients are very similar en each type of legs, showing a very similar increase of temperature with height.

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.495\textwidth,clip=]{figs/perfil_paper_ne_cr2082_up.eps}
\includegraphics[width=0.495\textwidth,clip=]{figs/perfil_paper_te_cr2082_up.eps}
\includegraphics[width=0.495\textwidth,clip=]{figs/perfil_paper_ne_cr2208_up.eps}
\includegraphics[width=0.495\textwidth,clip=]{figs/perfil_paper_te_cr2208_up.eps}
\caption{{Average fits to $\Ne(r)$ (left panels) and $T_e(r)$ (right panels) for legs of type I (red), II (violet), and III (green), for CR-2082 (top panels) and CR-2208 (bottom panels). Solid lines correspond to DEMT results while dashed lines correspond to AWSoM results.} \notebyalbert{(by Albert): The idea of this Figure is to show how DEMT and AWSoM compare AS A FUNCTION OF HEIGHT, in three distinct types of magnetic structures, and for both rotations.}}
\label{perfiles_promedio}
\end{center}
\end{figure}

Tu summary the statistical comparison, Figure \ref{histos_2082} and \ref{histos_2208} shows the statistical results of both rotations for the three types of field lines traced with DEMT (blue) and AWSoM results (red). From left to right the panels show the statistical distribution of $\NCB$, $\lN$ and $\aTm$. For all types of lines, Table \ref{tabla_comp} summarizes the median value of the statidtical distribution of each physical quantity derived from the analysis. The DEMT quantities are expressed as absoulte values, while for ASWSoM they are informed relative to the corresponding result for DEMT.

\diego{aca podria ir un pequeño parrafo que resuma las diferencias. Pero justo a continuacion empiezan las conclusiones y uno tiende a repetir, no se como encararlo para que no quede dos veces lo mismo y tan pegado uno de lo otro.}

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.31\textwidth,height=0.18\textwidth,clip=]{figs/histo_2082_demt_awsom_streamer_up_ne_1055.eps}
\includegraphics[width=0.31\textwidth,height=0.18\textwidth,clip=]{figs/histo_2082_demt_awsom_streamer_up_lambda_n.eps}
\includegraphics[width=0.31\textwidth,height=0.18\textwidth,clip=]{figs/histo_2082_demt_awsom_streamer_up_Tm.eps}
\includegraphics[width=0.31\textwidth,height=0.18\textwidth,clip=]{figs/histo_2082_demt_awsom_bound_up_ne_1055.eps}
\includegraphics[width=0.31\textwidth,height=0.18\textwidth,clip=]{figs/histo_2082_demt_awsom_bound_up_lambda_n.eps}
\includegraphics[width=0.31\textwidth,height=0.18\textwidth,clip=]{figs/histo_2082_demt_awsom_bound_up_Tm.eps}
\includegraphics[width=0.31\textwidth,height=0.18\textwidth,clip=]{figs/histo_2082_demt_awsom_CH_up_ne_1055.eps}
\includegraphics[width=0.31\textwidth,height=0.18\textwidth,clip=]{figs/histo_2082_demt_awsom_CH_up_lambda_n.eps}
\includegraphics[width=0.31\textwidth,height=0.18\textwidth,clip=]{figs/histo_2082_demt_awsom_CH_up_Tm.eps}
\caption{{Statistical distribution of DEMT (blue) and AWSoM (red) results traced along legs of type I, II and III (from top to bottom), as defined in Section 2.3. From left to right: electron density at the lowest coronal height of the AWSoM model $\Ne(r=1.055\,\mrsun)$, electron density scale height $\lN$, and leg-averaged electron temperature $\aTm$. In each panel the median value $m$ is indicated.}}
\label{histos_2082}
\end{center}
\end{figure}

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.31\textwidth,height=0.18\textwidth,clip=]{figs/histo_2208_demt_awsom_streamer_up_ne_1055.eps}
\includegraphics[width=0.31\textwidth,height=0.18\textwidth,clip=]{figs/histo_2208_demt_awsom_streamer_up_lambda_n.eps}
\includegraphics[width=0.31\textwidth,height=0.18\textwidth,clip=]{figs/histo_2208_demt_awsom_streamer_up_Tm.eps}\\
\includegraphics[width=0.31\textwidth,height=0.18\textwidth,clip=]{figs/histo_2208_demt_awsom_bound_up_ne_1055.eps}
\includegraphics[width=0.31\textwidth,height=0.18\textwidth,clip=]{figs/histo_2208_demt_awsom_bound_up_lambda_n.eps}
\includegraphics[width=0.31\textwidth,height=0.18\textwidth,clip=]{figs/histo_2208_demt_awsom_bound_up_Tm.eps}\\
\includegraphics[width=0.31\textwidth,height=0.18\textwidth,clip=]{figs/histo_2208_demt_awsom_CH_up_ne_1055.eps}
\includegraphics[width=0.31\textwidth,height=0.18\textwidth,clip=]{figs/histo_2208_demt_awsom_CH_up_lambda_n.eps}
\includegraphics[width=0.31\textwidth,height=0.18\textwidth,clip=]{figs/histo_2208_demt_awsom_CH_up_Tm.eps}
\caption{Same as Figure \ref{histos_2082} for CR-2208.}
\label{histos_2208}
\end{center}
\end{figure}

%This plot (E) has never been shown before and I believe it is a very nice result. There are of course previous studies on this, but this would be the first time this is evaluated with tomography. It is not the terminal speed, but it already shows the fast/slow components. We can show this, or we can go for more. Diego can QUANTIFY the correlation by carefully mapping all open field lines at 6 Rsun down to the DEMT region. He needs to code a bit more to achieve this goal, but we believe it is worth. With that tool done Diego can explore other terminal values of quantities of the AWSoM model and the DEMT results below (expansion factor maybe?). This will be nice to do also for the July-2-2019 Eclipse rotation that we will analyze after this paper.

\begin{table}
\begin{tabular}{l r@{.}l@{\hskip 0.05in} r@{\hskip 0.01in} r@{.}l  r@{.}l@{\hskip 0.05in} r@{\hskip 0.01in} r@{.}l r@{.}l@{\hskip 0.05in} r@{\hskip 0.01in} r@{.}l }
% r@{.}l@{\hskip 0.05in} r@{\hskip 0.01in} r@{.}l@{\hskip 0.01in} l}
\hline
Type    & \multicolumn{5}{c}{$\med(\NCB)$}             & \multicolumn{5}{c}{$\med(\lN)$} & \multicolumn{5}{c}{$\med(\avgTe)$} \\
       & \multicolumn{5}{c}{$[10^8\,{\rm cm}^{-3}]$}  & \multicolumn{5}{c}{$[{\rm 10}^{-2}\,\mrsun]$} & \multicolumn{5}{c}{$[\MK]$} \\
\hline
CR-2082\\
I  & 1&13 &(\Mi&11&9\%)  &   8&3 &(\Mi&~2&4\%) &   1&29 &(\Mi&10&8\%) \\
II  & 0&97 &(\Mi&22&5\%)  &   8&9 &(\Pl&~2&2\%) &   1&33 &(\Mi&~4&5\%) \\
III  & 0&66 &(\Mi&16&8\%)  &   7&7 &(\Mi&~9&1\%) &   0&98 &(\Mi&~9&2\%) \\
\hline          
CR-2208\\
I  & 0&97 &(\Pl&~9&3\%)  &   9&8 &(\Mi&~7&1\%) &   1&55 &(\Mi&14&2\%) \\
II  & 0&77 &(\Pl&11&7\%)  &  11&6 &(\Mi&17&2\%) &   1&55 &(\Mi&~9&7\%) \\
III  & 0&56 &(\Mi&~1&9\%)  &   8&6 &(\Mi&16&3\%) &   1&04 &(\Mi&~9&6\%) \\
\end{tabular}
\caption{Median value (indicated as ``Md'') of the statistical distribution of $\NCB$, $\lN$, and $\aTm$ for each coronal type of lines defined in Section \ref{trace}. DEMT values are expressed in absolute terms, while AWSoM results are informed as a percentual variation relative to the corresponding DEMT value. \albert{Propongo que para todas las cantidades absolutas informes siempre dos cifras significativas, y que todos los porcentajes los redondees a cero decimales (o sea, a veces una y a veces dos cifras significativas) y que no uses puntos para estos nunca. Ejemplo para CR-2082 $\NCB$ de Type-II: $1.0\,(-23\%)$. Ejemplo para CR-2208 $\lambda_N$ de Type-II: $12.0\,(-17\%)$. Ejemplo para CR-2208 $\aTm$ de Type-II: $1.6\,(-10\%)$.}}
\label{tabla_comp}
\end{table}

Figure \ref{perf_lon_vr} show the longitude-averaged AWSoM radial wind speed $V_r$ at $6\,\mrsun$, where all field lines are open. The heliocentric current sheet (HCS) location is indicated by the minimum of the speed curve.\\
\diego{can we say something without having traced the lines?} \notebyalbert{(by ALBERT) Todo el punto de la última Figura es precisamente decir algo. La comparación con la Figura 10 muestra una anti-correlación entre la densidad basal de DEMT y la velocidad terminal de AWSoM.}

For each rotation, everything to the south of the HCS maps down to the southern open region in the Figures \ref{perf_lat}, and the same for the northern hemisphere, showing an anti-correlation between $\Ne$ at low heights and high $V_r$ at $6\,\mrsun$.

\diego{Voy a intentar jugar con las lineas tipo 3 aumentando su cota inferior de latitud para ver si los histogramas de temp media dejan de ser tan dispersos y se vuelven mas sub-1MK. }

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.495\textwidth]{figs/Perfil_Vr_2082_5995.eps}
\includegraphics[width=0.495\textwidth]{figs/Perfil_Vr_2208_5995.eps}
\caption{{Longitude-averaged latitudinal dependence of the AWSoM model wind speed at $6.0\,\mrsun$ for CR2082 (left panel) and CR2208 (right panel).} \notebyalbert{(by Albert): Even if this is not yet the terminal speed, it is a pretty high height, and one can already see the transition from FAST to SLOW speed in the model, where the absolute minimum indicates the location of the HCS at this height. I believe it is interesting to compare these two curves with Figure 10. Note that while in this Figure all latitudes are magnetically open, in Figure 10 only the part at latitudes larger than the vertical black lines are open. Note that DEMT shows that in the open low corona $N_e$ decreases from the O/C boundary towards the poles, anti-correlating with the AWSoM terminal speed, which is nice. In order to quantify this anti-correlation we need to trace the terminal speed and the DEMT $N_e$, which we may need to do with the help of Chip and Nishtha. Chip, we can discuss this idea over Skype after Xmas if you are around.}}
\label{perf_lon_vr}
\end{center}
\end{figure}

\section{Discussion}\label{discu}

\noindent\notebyalbert{(by Albert) This section will discuss and summarize some main conclusions, possibly in three groups: a) DEMT analysis alone (including thermodynamical structure, up/down loops, energy considerations, etc.), b) Comparison AWSoM/DEMT (longitude-averaged latitudinal profiles, height-dependence profiles in different magnetic structures, etc), and c) Correlation betweem AWSoM terminal wind speed and DEMT basal density, I will expand on this later this week as this may require help from Michigan).}

\diego{NOTE: Here is a paragraph describing the empirical successes in the reconstruction of both rotations.}

*In comparing the DEMT results obtained for the two selected targets, it is important to bear in mind they rely on data provided by two different instruments: AIA and EUVI. In order to quantify the systematic difference of the DEMT products based on both instruments, \citet{nuevo_2015}, who were the first to apply DEMT to AIA data, analyzed a single target using both instruments independently. They found that while the density product is essentially equal, the temperature product of DEMT based on AIA data is systematically 8\% larger than the one based on EUVI data, i.e. $\Tm^{\rm(AIA)}/\Tm^{\rm(EUVI)} \approx 1.08$. Considering such correction, Figure \ref{histos_fulldemt} and Table \ref{tabla_demt} indicate that CR-2208 was $\approx 10-15\%$ hotter than CR-2082 throughout the streamer belt region. As for the electron density products, CR-2208 was found to be $\approx 15-20\%$ less dense than CR-2082 throughout the streamer belt region. These systematic differences are beyond the uncertainty level in the DEMT products due to systematic sources (radiometric calibration and tomographic regularization), that \citet{lloveras_2017} estimated to be $\lesssim 5\%$ and $\approx 2\%$ for the electron temperature and density, respectively.

*Coronal magnetic structures for which the temperature increases/decreases with height in the range of heights covered by DEMT, $1.02–1.22\,\mrsun$, have been dubbed as “up”/“down” loops by \citet{huang_2012} and \citet{nuevo_2013}, who first observed their presence by means of DEMT. As speculated by the authors of those works, loops of type down can be expected if the heating deposition is strongly confined near the coronal base of a magnetic loop. Down loops were first predicted by \citet{serio_1981}, and later on by \citet{aschwanden_2002}. In a recent study by \citet{schiff_2016}, down and up loops have been successfully reproduced by a numerical implementation of a 1D steady-state model that considers time-averaged heating rates.

We found down legs in the DEMT reconstructions mainly in the equatorial latitudes, being a greater quantity and better distribution along longitudes in CR-2082 than in CR-2208. \citet{nuevo_2013} showed a decrease in the amount of down loops as the solar activity increases, which is consistent with the results shown in this article.

*We can see, on type 0 loops, a median value of energy input flux ($\phi_h$) for CR-2082 twice bigger than CR-2208. In both rotations this magnitude has negative values, this can refer to some loops where loop-integrated radiative flux $\phi_r$) is too small in comparison with the larger gradients of negative conductive flux $\phi_c$), resulting in negative energy input flux.

This effect can be due a subestimation of the density present in the loop or the limited range of temperature used to reconstruct tomographic parameters, that can affect the compute of radiative power. Nevertheless, this population represent lower than the \diego{ALGÚN PORCENTAJE$\%$} of total population in both cases.

Consistent that the density in CR-2082 was higher, the $\phi_r$ was also higher. In both rotations the magnitudes in the temperature gradients are similar, but the temperatures in CR-2208 were consistently higher at all heights, resulting in a larger $\phi_c$ in module. This results are in agree with the observations of \citet{maccormack_2017}.

\diego{THE AR of CR-2082 wasn't well modeled by AWSoM.}

In both rotations, AWSoM and DEMT resutls shows same termodynamic structures along latitude and longitude with very comparable values between $1.055 - 1.25\,\mrsun$

Figure \ref{histos_2082} and Table \ref{tabla_comp} indicate that AWSoM reconstruction of CR-2082 was found to be less dense $\approx 11-22\%$ and colder $\approx 5-10\%$.

Both reconstructions of CR-2208 reconstructed the ARs with very similare values and in almost the same latitude-longitude location. Figure \ref{histos_2208} and Table \ref{tabla_comp} indicate that AWSoM reconstruction of CR-2208 was found to be denser $\approx 0-10\%$ and colder $\approx 10-15\%$.

CR-2208 was $\approx 10-15\%$ hotter than CR-2082 throughout the streamer belt region. As for the electron density products, CR-2208 was found to be $\approx 15-20\%$ less dense than CR-2082 throughout the streamer belt region. These systematic differences are beyond the uncertainty level in the DEMT products due to systematic sources (radiometric calibration and tomographic regularization), that \citet{lloveras_2017} estimated to be $\lesssim 5\%$ and $\approx 2\%$ for the electron temperature and density, respectively.

\diego{I would like to say something about FIgure 10. Maybe the weird tale in AWSoM Ne in the poles. Ideas?}\\

%\diego{We could say something about the anti correlation between speed and density and not get into the trace because issues do always appear and we don't have much time, and we all want to be happy in january. Unless the collaborators find easy to make this part.}

%% Figure 
%
% \begin{figure} 
% \centerline{\includegraphics[width=0.5\textwidth,clip=]{<fig.eps>}}
% \caption{}%\label{fig:?}
% \end{figure}

%% Table
%
% \begin{table}
% \caption{}%\label{tbl:?}
% \begin{tabular}{}     
% \hline
% \multicolumn{2}{c}{<>}
% <data>
% \hline
% \end{tabular}
% \end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Appendix
%
% \appendix

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Acknowledgements
%
% \begin{acks}
%
% \end{acks}

%%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Bibliography
%
% Using BibTeX
%
\bibliographystyle{spr-mp-sola}
\bibliography{solphys_dglloveras}  
%
% Without BibTeX 
% \begin{thebibliography}{}
% \bibitem[\protect\citeauthoryear{Author}{Year}]{key}
%   <bibliographical entry>
%
% \bibitem[\protect\citeauthoryear{}{}]{}
%   
%  
% \end{thebibliography}

\end{article} 
\end{document}