### 2020-02-21 response_to_review_SOLA-D-20-00015_v1.txt
..

Dear Editor,

Thank you very much for your email with the report on the first version of our manuscript SOLA-D-20-00015, titled "Thermodynamic Structure of the Solar Corona: Tomographic Reconstructions and MHD Modeling".

We deeply acknowledge the reviewer, whose report reflects her/his very careful and interested read, containing a series of relevant, useful, and constructive criticisms and suggestions.

Below, we paste the full review. Following each query we include a response, highlighting the changes in the manuscript, where all edited/new text is marked in RED COLOR fonts.

Looking forward to your response. Sincerely,
The Authors.

------------------------------begin of response------------------------------------------
COMMENTS FOR THE AUTHOR:

Review of Thermodynamic Structure of the Solar corona: tomographic reconstructions and MHD modeling

This is a well-structured and clearly-written paper, presenting a detailed comparison of tomographic reconstructions of the low solar corona (which gives density and temperature), and a global MHD model. The work is important and valuable to the field, and is suitable for publication in Solar Physics. There are no major scientific concerns. The paper should be ready for publication if the following, mostly minor, concerns are addressed:

Medium concerns
-----------------------------------------------------------------------------------------
1. All of the multi-panel figures that show longitude-latitude maps (e.g. figure 1, 2, 3 etc) are too small, it is very difficult to see any details. There is a lot of white space between the panels that can be removed, and this would be one way to instantly make the maps larger.

We fully  agree. We improved the use of space in Figures 1 through 3. We also changed colors in the latter (see also query #2 below). Let us know.

(ESTO ES TRABAJO PARA DIEGO & ALBERT, XTOOLS).

-----------------------------------------------------------------------------------------
2. Figure 3 and 11, bottom panels. I find it difficult to distinguish between the red and magenta colours. Can you replace magenta with a different colour, maybe green?

We fully agree. Actually this affects more figures: Figures 3, 4, 11, and 14. In all of these we made the changes: RED->ORANGE and MAGENTA->DarkRed. These changes emulate our previous choice of colors in Nuevo et al. 82013) and Lloveras et al. (2017). We like to keep all "up" structures in "warm" palette. We edited text accordingly all along the manuscript.

(ESTO ES TRABAJO PARA DIEGO, CON LA ASESORIA ARTISTICA DE FEDE y ALBERT).

-----------------------------------------------------------------------------------------
3. Page 14, paragraph starting at line 34. You discuss the latitudinal distribution of short and long loops, and state that the small legs mainly populate the equatorial latitudes, and longer legs at higher latitudes. But this is a skewed result - their spatial positioning was a criteria for their categorisation, so this should not be presented as a scientific result.

DIEGO, SUGERIMOS CON FEDE QUE REHAGAS COMO EXPERIMENTO LOS PANELES INFERIORES DE LA FIGURAS 3 CON OTRA DEFINICION DE TYPE 0, I Y II. SE QUE LO HICIMOS, PERO LO REVISITARIA Y SUGERIMOS AGREGAR UN GRAFICO DE HISTOGRAMA DE FOOTPOINT DE LAS CUATRO CLASES. FIJATE:

* TYPE-0:   CLOSED-SMALL-DOWN
* TYPE-I:   CLOSED-SMALL-UP
* TYPE-II:  CLOSED-LARGE-UP
* TYPE-III: TAL COMO ESTÁ (PARA QUE NO NOS ENSUCIE LA ESTADISTICA DE RESULTADOS DEMT LOS CHs DE LAT BAJA)

SI VEMOS QUE NO ES TAN SUCIO EL RESULTADO, PODEMOS DEJAR ESA SELEECCION YE INCLUIR QUIZA UN NUEVO GRAFICO: HISTOGRAMA DE FOOTPOINT DE LOS 4 TIPPOS. (UNA SOLA FIURA DE 4 COLORES) QUE SEAN LOS NUEVOS DOS PANELES INFERIORES DE LA fIGURA 3, QUE TENDRIA 6 PANELES EN TOTAL. LUEGO DECIDAMOS SI QUEDA COMO ESTA O USAMOS UNA NUEVA VERSION, Y EN CUALQUIERA DE AMBOS CASOS RETOCAMOS UN POCO EL TEXTO DE PAG 14 QUE QUEDO "MISSLEADING".

-----------------------------------------------------------------------------------------
4. Table 2 summarises the differences of the median values of the DEMT and MHD model. This is OK. However, there is an opportunity here to present a more detailed comparison, including in particular the mean (or median) absolute deviation (MAD) of values between DEMT and MHD. The MAD is a more stringent, voxel-by-voxel test of the comparison, which cannot be seen from the histogram comparisons (e.g. figure 12/13) nor the median values of Table 2.

REFERENCIA: https://en.wikipedia.org/wiki/Median_absolute_deviation

DIEGO, PROPONEMOS HACER ESTO CON FEDE.

DEFINAMOS EL CAMPO ESCALAR Xi = (X_MHD - X_tom) / X_tom, DONDE "X" ES CUALQUIER RESULTADO A LO LARGO DE LOOP (Ncb, LAMDA_n, ó <Tm>). LA TABLA 2 INFORMA LOS VALORES: "[Med(X_MHD) - Med(X_tom)]/Med(X_tom)".

PROPONEMOS DEJAR LOS HISTOGRAMAS DE LAS FIGS 12 Y 13 COMO ESTÁN, PERO EXPANDIR LA TABLA 2 A QUE INCLUYA DOS COLUMNAS MÁS INDICANDO ESTOS DOS INDICADORES DE CADA CAMPO "Xi":
* Med(Xi)  (debería aproximarse a los porcentuales y signos entre paréntesis que actualmente informa la Tabla 2).
* MAD = Med(Abs(Xi-Med(Xi))).

LUEGO EDITAMOS TABLA Y TEXTO DEL PAPER.

-----------------------------------------------------------------------------------------
5. You discuss the less reliable results from AIA in coronal holes. The very recent paper by Saqri et al (Solar Physics 295, 2020) is relevant here, and should be referenced.

YO PROPONGO HACER DOS O TRES EXPERIMENTOS DE CARACRTERIZACION Y RESOLUCION DEL PROBLEMA TESTEANDOLO A LAS TRES ALTURAS DEL PAPER. SI VEMOS QUE SALIO, POR QUE NO MEJORAR EL PAPER.

-----------------------------------------------------------------------------------------
6. The abstract can be shortened somewhat, reducing the number of introductory sentences and with a little less emphasis on the method and future improvements, and more on the science results. In particular: the results of the energy balance analysis and the estimate of energy flux at the coronal base, which are important.

(TOTALMENTE, HAGAMOS ESTO LUEGO DE EDITAR TODO).

-----------------------------------------------------------------------------------------
7. Page 5, paragraph beginning line 34. You briefly describe fitting the DEM to a Gaussian. I know this approach is covered in detail in Nuevo 2015, but I would like a little more detail here. You should mention that, in general, DEMs do not necessarily have a clean Gaussian profile and that more complicated inversions are needed (reference https://ui.adsabs.harvard.edu/abs/2012A%26A...539A.146H/abstract , https://ui.adsabs.harvard.edu/abs/2019SoPh..294..135M/abstract , https://ui.adsabs.harvard.edu/abs/2019SoPh..294..136P/abstract ). Then explain briefly why DEMT leads to cleaner, non-multithermal profiles (due to the 3D spatial reconstruction of the line of sight). I assume that this is why you can use a straightforward Gaussian fit? Also efficiency?

Thanks for this comment and the references. As the LDEM describes the temeprature distribution of the plasma in a specific voxel it does not deal with different large scale structures, as it usually it may be the case for the DEM describing the plasma alog a full line-of.sight (LOS). In that sense one can expect as you say, cleaner, profiles than those observed by DEM studies. Still, if one adds tbhe 335 A channel, which adds two sensitivity poits at ~ 1.0 MK and ~ 2.5 MK, one finds that a a bi-Gaussian LDEM model best predics all the tomographic emissivities. That is the main conclusion of Nuevo et. al. (2015). The interpretation in that case is that the unresolved independent strands threading the voxel can have two different characteristic temperatures. Nuevo et al. (2015). So in this sense, the plasma within the voxel can be multi-thermal. As for efficiency, the parametric technique is comparably efficient independently of the chosen model. We did read the  We expanded this discussion in the text.

-----------------------------------------------------------------------------------------
8. The introduction does not have many references to works beyond DEMT and AWSoM. The 2nd paragraph of the introduction offers a chance to remedy this, where you discuss the quiet-Sun corona. I know there are very few works that address the quiet corona in the context of DEM analysis, but you can reference https://www.aanda.org/articles/aa/full_html/2014/04/aa23054-13/aa23054-13.html. You can also mention https://iopscience.iop.org/article/10.1088/0004-637X/795/2/111 here (although you do reference this paper in more detail in the discussion). Another relevant paper which can be referenced, with DEM analysis of a large amount of quiet coronal data over long time periods, is https://advances.sciencemag.org/content/3/7/e1602056 .

ALBERT VA A ESCRIBIR UN POCO DE TEXTO.

LO QUE SIGUE DEBE SER TODO HECHO POR DIEGO, CONSULTANDO PRINCIPALMENTE A ALBERT. TODO CAMBIO DEBERÁ ESTAR DESTACADO CON \EDIT{} EN ROJO.
-----------------------------------------------------------------------------------------
9. Some sections can benefit from weeding out of repetitions. You tend to repeat statements regarding the method a lot. For example:
- Paragraph starting page 13, line 41. This has already been said in the method.
- Sentence Page 14, line 25 'For each leg…
- statement in parenthesis at page 15, line 24.
- Section 4, first two paragraphs.
- Page 26, line 26:
 'In the streamer belt region we combined the DEMT results and the AWSom magnetic model to infer the energy input phi at the coronal base required to maintain stable coronal loops. We found characteristic values in the range…'
Shorten to: 'The energy input phi at the coronal base, required to maintain stable coronal loops, is in the range…'
- Page 26, lines 38-43, between 'Results of the AWSoM….summarized as follows.' can be omitted.

-----------------------------------------------------------------------------------------
Minor grammatical improvements, typos, and removal of redundant statements
Page number, line number: 
3,36: quantitive analysis
3, 44: based on data
4, 25: absent => is this the right choice of word? I would think that the effect of fast dynamics is reduced, rather than being completely absent?
Equation 1 (and subsequent). You use LDEM and TRF. Is it more correct to use Greek symbols rather than acronyms in the equations? However, if you have used these acronyms in equations in previous studies, please leave as is.
6, 9: average absolute relative difference
6, 35: energy partitioning is improved
6, 46: broadened
6, 50 and 51: explain the subscripts for N and T
7, 39: waves is set to L
7, 41: radially stretched => is there a better description? 
7, 41: The spherical grid… has an adaptive grid that has fine resolution near the Sun, and increases…
7, 46: million cells on 6 x 
7, 49: The AWSoM simulation of the solar corona for…to the DEMT in the results section.
8, 23: products are traced
8, 33: For DEMT the data points used are 
8, 48: hereafter, with del_r=e_r.del being the radial derivative operator, and e_r is the 
9, 2: to directly compare the observed
9, 11: , and later by Aschwanden
9, 17: and the heliocentric height r. We then
9, 19: the goodness of fit  we perform
9, 21: uncertainty of the DEMT, selecting legs
9, 24: linear fit to temperature is weak, are excluded
9, 28: according to del_r
9, 31: quality of the fit considers
9, 33: In summary, a selected leg must meet three conditions:
9, 40: temperature and height points must meet
9, 48: four different types:
10, 12: type 0 to II allows a study of the streamer belt in three distinct layers at increasing height.
10, 25: model, as described
10, 35: transverse => cross-sectional?
11, 44: Figures 1 and 2 show …. (omit 'As an example', not needed)
12, 35: Rotation CR-2082 has two small… (and same in following sentence)
13: 34: by Equation 4 show that the agreement… (omit 'in most of the coronal volume of the DEMT grid', not needed)
13: 36: A similar result was
15, 19: to this point in the Conclusions.
15, 28: Caption to table 1. Informed => given? Also explain that the CR-2208 values are given in parenthesis.
15, 42: Table 1 summarises the results of the…
16, 51: , the temperature based on AIA data
17, 45: Considering this error, Figure 5
18, 8: legs, each loop was
18, 20: characterised    by temperatures that are are progressively larger, and farther from T_c, for both rotations.
18, 33: smallest sized loops of type 0, and it is likely due to emission out of the instrumental sensitivity range.
19, 17: axisymmetric, similar to the tomographic results. (In fact, you should avoid using 'tomographic model' since it is not a model, this happens a few times in the text).
19, 45: a visual comparison reveals similar
19, 50: Such a comparison…, with the top
20, 30: are worth highlighting.
20, 43: boundary. Lastly, the DEMT… ('note that' is usually not needed, you use quite a lot in text)
20, 47: For comparison, B_r in the CHs increases from
22, 42: tomografic => tomographic
22, 48: This figure can be compared
22, 49: …DEMT results. The AWSoM maps are more densely populated
22, 50: having spatially smoother distributions of electron density and temperature than those of DEMT.
23, 38: variation relative to the DEMT. (Then omit 'The following major results can be drawn', not needed).
25, 47: Omit 'Concerning the DEMT reconstructions, Figure 3 shows that, for both rotations, …'  Start paragraph with 'Magnetic field lines of type…'
25, 50: Omit 'Figure 5 and Table 1 show in 3D quantitive detail how…', and start sentence with 'These progressively outer'
26, 12: such comparison provided a consistency check. (And omit 'for the changes implemented in the DEMT technique.')
26, 20: less reliable for CHs.
26, 24: algorithm to deal
27, 12: This is likely the cause of a more accurate match to the…
27, 14: Omit first sentence of paragraph, not needed.
27, 30: in the range of
27, 33: This will be investigated in

-----------------------------------------------end of response---------------------------