### 2020-03-10 response_to_review_SOLA-D-20-00015_v1.txt
Albert está trabajando en esto, no tocar todavía, gracias.

Dear Editor,

Thank you very much for your email with the report on the first version of our manuscript ID SOLA-D-20-00015, titled "Thermodynamic Structure of the Solar Corona: Tomographic Reconstructions and MHD Modeling".

We deeply acknowledge the reviewer, whose report reflects her/his very careful and interested read. We found all queries very relevant, with useful and constructive criticisms and suggestions. Below, we copy/paste the full review. Following each query (Q) we include a corresponding answer (A), commenting the changes in the manuscript, in which all edited/new text is highlighted in red color font.

We addressed all points, as a result of which we believe the manuscript improved in both content and clarity. In particular, a major improvement in this new version of the draft relates to Q5 below.

Looking forward to your response. Sincerely,
The Authors.

------------------------------beginning of review and response---------------------------
COMMENTS FOR THE AUTHOR:

Review of Thermodynamic Structure of the Solar corona: tomographic reconstructions and MHD modeling

This is a well-structured and clearly-written paper, presenting a detailed comparison of tomographic reconstructions of the low solar corona (which gives density and temperature), and a global MHD model. The work is important and valuable to the field, and is suitable for publication in Solar Physics. There are no major scientific concerns. The paper should be ready for publication if the following, mostly minor, concerns are addressed:

Medium concerns

-----------------------------------------------------------------------------------------
Q1. All of the multi-panel figures that show longitude-latitude maps (e.g. figure 1, 2, 3 etc) are too small, it is very difficult to see any details. There is a lot of white space between the panels that can be removed, and this would be one way to instantly make the maps larger.

A1. We agree. We improved the use of space in the long-lat maps in Figures 1, 2, 8 and 9, while the one in Figure 3 remains the same (we believe those are fine).

-----------------------------------------------------------------------------------------
Q2. Figure 3 and 11, bottom panels. I find it difficult to distinguish between the red and magenta colours. Can you replace magenta with a different colour, maybe green?

A2. We agree. Actually this affects all following Figures: 3, 4, 11, and 14. In those, we applied the following changes: Red->Orange and Magenta->Red. These emulate our previous choice of colors in Nuevo et al. (2013) and Lloveras et al. (2017). We like to keep all "up" structures in a "warm" palette. We edited text accordingly all along the manuscript. We believe the colors are now clearly distinguishable in all figures.

-----------------------------------------------------------------------------------------
Q3. Page 14, paragraph starting at line 34. You discuss the latitudinal distribution of short and long loops, and state that the small legs mainly populate the equatorial latitudes, and longer legs at higher latitudes. But this is a skewed result - their spatial positioning was a criteria for their categorisation, so this should not be presented as a scientific result.

A3. Several aspects are involved here, the discussion in the paragraph you indicate strongly relates on that in the final paragraph of Section 2.3. In reading them we find that these texts were confusing as they were, so we edited both. Actually, we do not say that *small* legs populate the equatorial latitudes and *large* ones the mid latitudes. We say that if one classifies the *small-closed* loops in DOWN and UP, the former peak near the equator and the latter peak near mid latitudes (at least for the deep minimum rotation CR-2082). The classification of legs in all four classes in Section 2.3 includes limits for the latitude of their footpoints, to distinguish different layers of the streamer belt (legs of type 0, I and II), as well as to separate field lines of the high latitude CHs (legs of type III). We included two new panels in Figure 3, showing frequency histograms of the latitude of the four types of legs at 1.105 Rsun. Within the range Abs(latitude) < 50 deg, and for the solar minimum target CR-2082, down-legs are peaked near the equator while up-legs peak at mid-latitudes. This is less obvious for CR-2208, although it is clear that down-legs peak in the core of the streamer belt. If one analyses Abs(latitude) > 50 deg it turns out that small-down legs are scarce and there is an extra population of small-up legs (not nearly as much as in the peaks seen at mid lats. Should have we not included the limit of Abs(latitude) < 50 deg for closed-small down and up legs (i.e., legs of type 0 and I), their histograms would still peak at around the same latitudes. This can be seen in Figure 7 of Nuevo et al. 92013), were they analyzed the small closed up and down loops for two rotations of solar minimum, without placing any limits in the latitude of the footpoints. As our main objective here is to be able to analyze progressively outer layers of the streamer belt, we include these limits in the latitude of the footpoints. We edited in full the final paragraph of Section 2.3, as well as the paragraph of Section 3.1 you point out (now split in two, shorter paragraphs). We trust the new graphs and discussions are clear now. Let us know.

-----------------------------------------------------------------------------------------
Q4. Table 2 summarises the differences of the median values of the DEMT and MHD model. This is OK. However, there is an opportunity here to present a more detailed comparison, including in particular the mean (or median) absolute deviation (MAD) of values between DEMT and MHD. The MAD is a more stringent, voxel-by-voxel test of the comparison, which cannot be seen from the histogram comparisons (e.g. figure 12/13) nor the median values of Table 2.

A4. First of all we note that, as a result of the discussion in Q5 below, a few numbers in Table 2 were updated (those corresponding to type-III legs for CR-2208). Now, when selecting field lines for analysis of both the DEMT and the AWSoM models, the classification of types of legs introduced in section 2.3 is used for each model separately. As a result, as shown by comparing the middle panels of Figure 3 and the bottom panels of Figure 11, and as discussed in the corresponding texts of the manuscript, the selection of field lines for DEMT is not the same. In comparing those Figures each color indicates comparable regions between the two analysis. It is so that Figures 12 and 13, and Table 2, which show DEMT-versus-AWSoM comparison for each rotation, are not done line-by-line but rather region-by-region (say: orange, red and cyan). The suggestion of calculating the MAD would require selecting the same field lines for both models, which is beyond the scope of the present comparative analysis. Informing the MAD is a good idea, being a precise measure of how the comparison between both models spreads. We will certainly implement such an analysis in a next effort. In the present draft, we trust the current comparison (Figures 12 and 13, and Table 2) provide are pretty informative summary of how both models compare. Let us know what you think.

-----------------------------------------------------------------------------------------
Q5. You discuss the less reliable results from AIA in coronal holes. The very recent paper by Saqri et al (Solar Physics 295, 2020) is relevant here, and should be referenced.

A5. Thanks for the reference, which is certainly relevant and now mentioned in the paper, as discussed below. We are not sure stray light contamination was actually the reason of the reported issue, which is now solved.

The discussion section of v1 of the draf include this sentence (now deleted): "The source of this distinct performance of DEMT when applied to EUVI or AIA data is related to the TRFs of the respective filter sets. We are currently implementing an improved LDEM determination algorithm to deal with this issue that will be used in a follow up article." We have successfully implemented the improved version of the code, as a result of which this problem has been solved. The issue was related to the algorithm that solves the optimization problem posed in each voxel to find the LDEM. When based on AIA data, the previous algorithm tended to converge to a non-optimal (local minimum of the objective function) in specific voxels located in the CHs. We addressed the problem by exploring multiple starting points when solving the optimization problem in cases where the initial solution was not good enough. In this draft, this improved algorithm affects results concerning type III legs for CR2208. As a result, the following Figures have been updated: 2, 3, 4, 5, 10, 13 and 14. You will notice all these Figures are just like the original ones, except the problem in the CHs of CR-2208 has been solved. Text throughout the manuscript was edited accordingly, in the following locations of v2 of the draft:
 - Red text in Pg 12 and Pg 13, and the red rext in Pg 15 informing characteristic values of Grad_r(Tm).
 - Numerical values concerning legs of type-III for CR-2208 in Tables 1 and 2.
 - Most of the third paragraph of the Conclusion section in V1 was devoted to this issue, so it was eliminated. The remaining text is now the end of the second paragraph of that section (highlighted in red color).

Concerning the effect of stray-light contamination of EUV images on DEMT products, our current data processing pipeline for EUVI data includes the deconvolution of the PSF, carefully determined by Shearer et al. (ApJL, 2012). We have not yet included such a procedure for AIA data, as we were not aware of a reliable determination of their PSF. Shearer et al. (2012) showed that the signal in the core of CHs may be strongly affected by stray light. Now, the effect in tomography turns out to be milder, due to temporal and spatial binning of the images. In the case of EUVI data, Lloveras et al. (2017) performed a careful analysis of effect of stray-light removal in the resulting DEMT products. PSF deconvolution basically increases the contrast. As a result, the off-limb radial brightness in EUV images becomes somewhat steeper, which has an impact in tomographic reconstructions. Specifically, Lloveras et al. (2017) -see the end of their section 3.6- found that the coronal base density increases a bit (5-15%) and the density scale height decreases a bit (5-10%), while the electron temperature remains largely unchanged. We understand that stray light correction is less serious for AIA data. Again, it may be strong in specific pixels of AIA images and yet the final effect in tomographic products may be mild. Still, we do want to explore the effect in DEMT of stray light removal in AIA data. It will be in particular interesting to see if stray-light removal helps the algorithm that finds de LDEM in voxels of the CHs. We acknowledge the reference on AIA data, and will investigate this in the future. We included a comment on this topic, as well as the reference by Saqri et.al., at the beginning of Section 2.1 (red text).

-----------------------------------------------------------------------------------------
Q6. The abstract can be shortened somewhat, reducing the number of introductory sentences and with a little less emphasis on the method and future improvements, and more on the science results. In particular: the results of the energy balance analysis and the estimate of energy flux at the coronal base, which are important.

A6. We agree. We edited the Abstract. Let us know what you think.

-- FALTA ESTO --

-----------------------------------------------------------------------------------------
Q7. Page 5, paragraph beginning line 34. You briefly describe fitting the DEM to a Gaussian. I know this approach is covered in detail in Nuevo 2015, but I would like a little more detail here. You should mention that, in general, DEMs do not necessarily have a clean Gaussian profile and that more complicated inversions are needed (reference https://ui.adsabs.harvard.edu/abs/2012A%26A...539A.146H/abstract , https://ui.adsabs.harvard.edu/abs/2019SoPh..294..135M/abstract , https://ui.adsabs.harvard.edu/abs/2019SoPh..294..136P/abstract). Then explain briefly why DEMT leads to cleaner, non-multithermal profiles (due to the 3D spatial reconstruction of the line of sight). I assume that this is why you can use a straightforward Gaussian fit? Also efficiency?

A7. Thanks for this comment and the references. We were aware of those, discussing the first one in Lloveras et.al. (2017). Concerning the other two, we are in fact interested in using the SITES code in the determination of the LDEM at each voxel, and make comparisons with those provide by our current parametric model code. As for the reason of LDEMs being "cleaner" than DEMs, you are right. As the LDEM describes the temperature distribution of the plasma in a specific voxel it does not deal with different large scale structures, as it may be the case for the DEM describing the plasma along a full line-of-sight (LOS). In that sense one can expect as you say, cleaner, profiles than those observed by DEM studies. Even so, if one includes in the analysis data from the 335 A band, which adds two sensitivity points at ~ 1.0 MK and ~ 2.5 MK, one finds that a a bi-Gaussian LDEM model best predicts all the tomographic emissivities. That is the main conclusion of Nuevo et. al. (2015). The interpretation in that case is that the unresolved independent strands threading the voxel can have two different characteristic temperatures. In this sense, the plasma within the voxel can be multi-thermal. As for efficiency, the parametric technique is comparably efficient independently of the chosen model. We included a new paragraph (red text) between Equations (1) and (2), discussing this point and including the references above and others as well.

-----------------------------------------------------------------------------------------
Q8. The introduction does not have many references to works beyond DEMT and AWSoM. The 2nd paragraph of the introduction offers a chance to remedy this, where you discuss the quiet-Sun corona. I know there are very few works that address the quiet corona in the context of DEM analysis, but you can reference https://www.aanda.org/articles/aa/full_html/2014/04/aa23054-13/aa23054-13.html. You can also mention https://iopscience.iop.org/article/10.1088/0004-637X/795/2/111 here (although you do reference this paper in more detail in the discussion). Another relevant paper which can be referenced, with DEM analysis of a large amount of quiet coronal data over long time periods, is https://advances.sciencemag.org/content/3/7/e1602056 .

A8. ALBERT VA A ESCRIBIR UN POCO DE TEXTO. -- FALTA ESTO --

LO QUE SIGUE DEBE SER TODO HECHO POR DIEGO, CONSULTANDO PRINCIPALMENTE A ALBERT. TODO CAMBIO DEBERÁ ESTAR DESTACADO CON \EDIT{} EN ROJO.

-----------------------------------------------------------------------------------------
Q9. Some sections can benefit from weeding out of repetitions. You tend to repeat statements regarding the method a lot. For example:
- Paragraph starting page 13, line 41. This has already been said in the method.
- Sentence Page 14, line 25 'For each leg…
- statement in parenthesis at page 15, line 24.
- Section 4, first two paragraphs.
- Page 26, line 26:
 'In the streamer belt region we combined the DEMT results and the AWSom magnetic model to infer the energy input phi at the coronal base required to maintain stable coronal loops. We found characteristic values in the range…'
Shorten to: 'The energy input phi at the coronal base, required to maintain stable coronal loops, is in the range…'
- Page 26, lines 38-43, between 'Results of the AWSoM….summarized as follows.' can be omitted.

A9. Thanks for pointing this out. We have improved the text in the following parts:

-----------------------------------------------------------------------------------------
Q10. Minor grammatical improvements, typos, and removal of redundant statements

A10. Thanks very much for this detailed corrections. We have addressed them all.

Page number, line number: 
3,36: quantitive analysis
3, 44: based on data
4, 25: absent => is this the right choice of word? I would think that the effect of fast dynamics is reduced, rather than being completely absent?
Equation 1 (and subsequent). You use LDEM and TRF. Is it more correct to use Greek symbols rather than acronyms in the equations? However, if you have used these acronyms in equations in previous studies, please leave as is.
6, 9: average absolute relative difference
6, 35: energy partitioning is improved
6, 46: broadened
6, 50 and 51: explain the subscripts for N and T
7, 39: waves is set to L
7, 41: radially stretched => is there a better description? 
7, 41: The spherical grid… has an adaptive grid that has fine resolution near the Sun, and increases…
7, 46: million cells on 6 x 
7, 49: The AWSoM simulation of the solar corona for…to the DEMT in the results section.
8, 23: products are traced
8, 33: For DEMT the data points used are 
8, 48: hereafter, with del_r=e_r.del being the radial derivative operator, and e_r is the 
9, 2: to directly compare the observed
9, 11: , and later by Aschwanden
9, 17: and the heliocentric height r. We then
9, 19: the goodness of fit  we perform
9, 21: uncertainty of the DEMT, selecting legs
9, 24: linear fit to temperature is weak, are excluded
9, 28: according to del_r
9, 31: quality of the fit considers
9, 33: In summary, a selected leg must meet three conditions:
9, 40: temperature and height points must meet
9, 48: four different types:
10, 12: type 0 to II allows a study of the streamer belt in three distinct layers at increasing height.
10, 25: model, as described
10, 35: transverse => cross-sectional?
11, 44: Figures 1 and 2 show …. (omit 'As an example', not needed)
12, 35: Rotation CR-2082 has two small… (and same in following sentence)
13: 34: by Equation 4 show that the agreement… (omit 'in most of the coronal volume of the DEMT grid', not needed)
13: 36: A similar result was
15, 19: to this point in the Conclusions.
15, 28: Caption to table 1. Informed => given? Also explain that the CR-2208 values are given in parenthesis.
15, 42: Table 1 summarises the results of the…
16, 51: , the temperature based on AIA data
17, 45: Considering this error, Figure 5
18, 8: legs, each loop was
18, 20: characterised    by temperatures that are are progressively larger, and farther from T_c, for both rotations.
18, 33: smallest sized loops of type 0, and it is likely due to emission out of the instrumental sensitivity range.
19, 17: axisymmetric, similar to the tomographic results. (In fact, you should avoid using 'tomographic model' since it is not a model, this happens a few times in the text).
19, 45: a visual comparison reveals similar
19, 50: Such a comparison…, with the top
20, 30: are worth highlighting.
20, 43: boundary. Lastly, the DEMT… ('note that' is usually not needed, you use quite a lot in text)
20, 47: For comparison, B_r in the CHs increases from
22, 42: tomografic => tomographic
22, 48: This figure can be compared
22, 49: …DEMT results. The AWSoM maps are more densely populated
22, 50: having spatially smoother distributions of electron density and temperature than those of DEMT.
23, 38: variation relative to the DEMT. (Then omit 'The following major results can be drawn', not needed).
25, 47: Omit 'Concerning the DEMT reconstructions, Figure 3 shows that, for both rotations, …'  Start paragraph with 'Magnetic field lines of type…'
25, 50: Omit 'Figure 5 and Table 1 show in 3D quantitive detail how…', and start sentence with 'These progressively outer'
26, 12: such comparison provided a consistency check. (And omit 'for the changes implemented in the DEMT technique.')
26, 20: less reliable for CHs.
26, 24: algorithm to deal
27, 12: This is likely the cause of a more accurate match to the…
27, 14: Omit first sentence of paragraph, not needed.
27, 30: in the range of
27, 33: This will be investigated in

-----------------------------------------------end of response---------------------------